---
title: "data exploration for house prices dataset"
output: html_document
date: "2025-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the required packages for this exploration. The code should run and
install packages if they are not installed on the local device

```{r}
requiredPackages = c('here', 'tidyverse', 'dplyr', 'data.table', 'ggplot2', 'rstatix', 'missForest', 'Metrics', 'glmnet', 'xgboost', 'mice', 'furrr', 'ufs', 'forcats', 'randomForest', 'future.apply', 'recipes', 'caret', 'broom', 'DescTools', 'infotheo', 'tibble', 'car', 'stats', 'poolr', 'keras', 'tensorflow', 'yardstick', 'h2o', 'ranger', 'lightgbm', 'Matrix')
for (p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p, character.only = TRUE)}
```

Here we're loading in the data required for this project

```{r}
setwd('C:/Users/marnt/Desktop/kaggle/house prices')
here() #Remember to reset the working directory as appropriate
train_csv <-as.data.frame(read.csv('train.csv'))
test_csv <- as.data.frame(read.csv('test.csv'))

```

We'll need to check the quality of the data before we can start predicting the sale prices. First check is to see where data is missing

```{r}
na_count_train <-sapply(train_csv, function(y) sum(length(which(is.na(y)))))
na_count_test <-sapply(test_csv, function(y) sum(length(which(is.na(y)))))
```

``` {r}
na_count_train
```
Missing data present in LotFrontage, Alley, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, FireplaceQu, GarageType, GarageYrBlt, GarageQual, GarageCond, PoolQC, Fence and MiscFeature fields.
We will need to further examine the data to decide whether or not to impute these missing values in some way

As a starting point, with missing values in the majority of entries for Alley, PoolQC, Fence, MiscFeature and FireplaceQu, we'll first investigate these features to determine whether they should be removed.


```{r}
train_csv %>% count(Alley)
train_csv %>% count(PoolQC)
train_csv %>% count(Fence)
train_csv %>% count(MiscFeature)
train_csv %>% count(FireplaceQu)
```
Missing values in FireplaceQu correspond to the property not having a fireplace, and we can impute missing values here easily. We will insert a new category to show that the properties do not have a fireplace. Imputing the missing values in other fields is similar in that we are assuming a missing value indicates the feature not being present on the property. This system applies similarly to the missing values in the fields about basements, masonry, and garages, although we will use a slightly different approach for those fields.

Note that there are two entries here in the basement field where information is missing for in just one basement field (BsmtExposure and BsmtFinType2) and not the other basement fields, which will be imputed by other methods.

```{r}
missing_cols_categorical <- c('FireplaceQu', 'Alley', 'PoolQC', 'Fence', 'MiscFeature', 'MasVnrType', 'GarageQual', 'GarageFinish', 'GarageCond', 'GarageType')
missing_cols_numerical <- c('MasVnrArea', 'GarageYrBlt')
for (col in missing_cols_categorical){
  train_csv[[col]][is.na(train_csv[[col]])] <- 'Missing'
}
for (col in missing_cols_numerical){
  train_csv[[col]][is.na(train_csv[[col]])] <- 0
}

bsmt_cols <- c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2')
all_missing <- apply(train_csv[, bsmt_cols], 1, function(row) all(is.na(row)))
train_csv[all_missing, bsmt_cols] <- 'Missing'
```
Before we start imputing values for variables which are missing data, like LotFrontage, and those associated with other house features, we'll need to further examine other features in the data.

Some variables are listed as numerical, but are in fact categorical and should be treated as such. These include the month that the property was sold (MoSold), the year the property was sold (YrSold) and the Subclass of the building (MSSubClass).

```{r}
train_csv$MoSold = as.factor(train_csv$MoSold)
train_csv$YrSold = as.factor(train_csv$YrSold)
train_csv$MSSubClass = as.factor(train_csv$MSSubClass)
```

This brings up another problem, some of the categorical variables in the dataset are ordinal and some are nominal, and they should be encoded as such to avoid problems further down the imputation pipeline. Some features (BsmtFinType2, BsmtExposure) have missing data that will be imputed so we will classify them as ordered after the imputation.

```{r}
#First, make everything that's text into a factor
character_cols <- lapply(train_csv, class) == 'character'
train_csv[,character_cols] <- lapply(train_csv[,character_cols], as.factor)

#Next, make sure that everything that has a scale (i.e. factors that are ordinal) has that scale defined in order
factor_vars <- lapply(train_csv, class) == 'factor'

#These are the ordinal factors that use the same scale, the other 7 need to be set individually
ordinal_vars <- c('ExterQual', 'ExterCond', 'BsmtQual', 'BsmtCond', 'HeatingQC', 'KitchenQual', 'FireplaceQu', 'GarageQual', 'GarageCond', 'PoolQC')                      
quality_levels <- c('Missing', 'Po', 'Fa', 'TA', 'Gd', 'Ex')
for (var in ordinal_vars){
  train_csv[[var]] <- factor(train_csv[[var]], levels = quality_levels, ordered = TRUE) 
}

#These are the other ordinal variables with their unique scales
BsmtFinType_levels <- c('Missing', 'Unf', 'LwQ', 'Rec', 'BLQ', 'ALQ', 'GLQ')
train_csv$BsmtFinType1 <- factor(train_csv$BsmtFinType1, levels = BsmtFinType_levels, ordered = TRUE)
Functional_levels <- c('Sal','Sev', 'Maj2', 'Maj1', 'Mod', 'Min2', 'Min1', 'Typ')
train_csv$Functional <- factor(train_csv$Functional, levels = Functional_levels, ordered = TRUE)
GarageFinish_levels <- c('Missing', 'Unf', 'RFn', 'Fin')
train_csv$GarageFinish <- factor(train_csv$GarageFinish, levels = GarageFinish_levels, ordered = TRUE)
Fence_levels <- c('Missing', 'MnWw', 'GdWo', 'MnPrv', 'GdPrv')
train_csv$Fence <- factor(train_csv$Fence, levels = Fence_levels, ordered = TRUE)
Slope_levels <- c('Sev', 'Mod', 'Gtl')
train_csv$LandSlope <- factor(train_csv$LandSlope, levels = Slope_levels, ordered = TRUE)
Paved_levels <- c('N', 'P', 'Y')
train_csv$PavedDrive <- factor(train_csv$PavedDrive, levels = Paved_levels, ordered = TRUE)
```

Before we go ahead and predict our missing values, we'll first try and examine the variables in the dataset and see which variables are strongly associated or un-associated with our outcome variables. We'll use Cramers V test for categorical variables and ANOVA tests for numerical variables.


```{r}
feature_usefulness <- function(df, target_var) {
  target <- df[[target_var]]
  
  # Drop NAs in target for association calculation
  df <- df[!is.na(target), ]
  target <- droplevels(df[[target_var]])
  
  results <- data.frame(
    Variable = character(),
    Type = character(),
    Measure = character(),
    Value = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (v in setdiff(names(df), target_var)) {
    # Skip variables that are entirely missing or constant
    if (all(is.na(df[[v]])) || length(unique(df[[v]])) < 2) next
    
    if (is.factor(df[[v]])) {
      # Categorical predictor — use Cramér’s V
      tbl <- table(df[[v]], target)
      val <- tryCatch(DescTools::CramerV(tbl, bias.correct = TRUE), error = function(e) NA)
      results <- rbind(results, data.frame(Variable = v, Type = "Categorical", Measure = "CramerV", Value = val))
      
    } else if (is.numeric(df[[v]])) {
      # Numeric predictor — use correlation ratio (η²)
      formula <- as.formula(paste(v, "~", target_var))
      val <- tryCatch({
        model <- aov(formula, data = df)
        summary(model)[[1]][["Sum Sq"]][1] / sum(summary(model)[[1]][["Sum Sq"]])
      }, error = function(e) NA)
      results <- rbind(results, data.frame(Variable = v, Type = "Numeric", Measure = "EtaSquared", Value = val))
    }
  }
  
  results <- results %>%
    arrange(desc(Value))
  
  return(results)
}

usefulness_tbl <- feature_usefulness(train_csv, target_var = "BsmtExposure")
significant_predictors <- usefulness_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
```

Now that we've got most of the missing categorical data appropriately assigned and imputed, we need to deal with the remaining missing values. Accomplishing this task will involve testing different methods to see which method provides the best results for imputation, as assessed by different methods (Accuracy of predictions, F1 Score). This function does several things that I'll describe briefly. First, there are a few safety checks that are completed as we select our significantly associated predictors (although this step is really more about removing very unassociated predictors), before a masked set of data is taken from the dataset and used to split the data into training / testing. We impute the values missing from the categories not being tested for (at this point LotFrontage, and the other categorical variables that are being imputed).Then with our cleaned data, we predict! As a baseline, we're using the mode values from the category as the outcome, which given the group sizes in these categories isn't a very bad idea. For two comparison methods, we'll use kNN and Random Forest classification models.

```{r}
evaluate_categorical_imputation <- function(df,
                                            target_var,
                                            predictor_vars = NULL,
                                            mask_frac = 0.20,
                                            k_knn = 5,
                                            rf_ntree = 200,
                                            seed = 123,
                                            max_predictor_missing_frac = 0.5) {
  set.seed(seed)
  if (!target_var %in% names(df)) stop("target_var not found in dataframe")
  # ensure target is factor
  if (!is.factor(df[[target_var]])) df[[target_var]] <- as.factor(df[[target_var]])
  
  results_tbl <- feature_usefulness(train_csv, target_var = target_var)
  significant_predictors <- results_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
  
  
  # choose predictors if not provided: all columns except target
  if (is.null(predictor_vars)) {
    predictor_vars <- significant_predictors$Variable
  }
  
  # Filter out predictors that are almost entirely missing
  predictor_missing_frac <- sapply(df[predictor_vars], function(x) mean(is.na(x)))
  keep_preds <- names(predictor_missing_frac)[predictor_missing_frac <= max_predictor_missing_frac]
  if (length(keep_preds) == 0) stop("No predictors with acceptable missingness found")
  predictor_vars <- keep_preds
  
  # Split known / missing target rows
  known_idx <- which(!is.na(df[[target_var]]))
  missing_target_idx <- which(is.na(df[[target_var]]))
  if (length(known_idx) < 10) stop("Too few known target rows to evaluate reliably")
  
  # Mask a random subset of known targets for evaluation
  mask_n <- max(1, round(length(known_idx) * mask_frac))
  mask_idx <- sample(known_idx, mask_n)
  df_masked <- df
  true_values <- df[[target_var]][mask_idx]
  df_masked[[target_var]][mask_idx] <- NA
  
  # Build training and testframes:
  # training rows = rows with target known (after masking)
  train_df <- df_masked %>% filter(!is.na(.data[[target_var]])) %>% select(all_of(c(target_var, predictor_vars)))
  test_df  <- df_masked[mask_idx, , drop = FALSE] %>% select(all_of(c(target_var, predictor_vars)))
  
  # Apply simple imputation to predictors using only training distribution for numerics:
  # For numeric medians we compute medians on training data and apply to both train/test,
  # For categorical we map NAs to "Missing".
  # Compute numeric medians from training.
  numeric_preds <- predictor_vars[sapply(train_df[predictor_vars], is.numeric)]
  medians <- sapply(train_df[numeric_preds], function(x) median(x, na.rm = TRUE))
  for (col in numeric_preds){
  train_df[[col]][is.na(train_df[[col]])] <- medians[[col]]
  test_df[[col]][is.na(test_df[[col]])] <- medians[[col]]
  }
  
  factor_preds <- setdiff(predictor_vars, numeric_preds)
  for (col in factor_preds){
    if(NA %in% train_df[[col]]){
      levs <- c(levels(train_df[[col]]), 'MissingVal')
      train_df[[col]] <- factor(train_df[[col]], levels = levs)
      test_df[[col]] <- factor(test_df[[col]], levels = levs)
    }
    train_df[[col]][is.na(train_df[[col]])] <- 'MissingVal'
  
    if(NA %in% test_df[[col]]){
      levs <- c(levels(test_df[[col]]), 'MissingVal')
      train_df[[col]] <- factor(train_df[[col]], levels = levs)
      test_df[[col]] <- factor(test_df[[col]], levels = levs)
    }
    test_df[[col]][is.na(test_df[[col]])] <- 'MissingVal'
  }
  
  # If medians are NA (all missing), set to 0
  medians[is.na(medians)] <- 0
  
  # Ensure target in train is factor and levels set
  train_df[[target_var]] <- factor(train_df[[target_var]])
  target_levels <- levels(train_df[[target_var]])
  
  # ---- 1) Baseline: Mode ----
  mode_value <- names(sort(table(train_df[[target_var]]), decreasing = TRUE))[1]
  pred_mode <- factor(rep(mode_value, nrow(test_df)), levels = target_levels)
  acc_mode <- mean(pred_mode == true_values, na.rm = TRUE)
  conf_mode <- table(Predicted = pred_mode, True = true_values)
  
  # ---- 2) kNN ----
  # Build dummy variables using model.matrix on combined train+test so columns match
  combined_preds <- bind_rows(
    train_df %>% select(all_of(predictor_vars)),
    test_df %>% select(all_of(predictor_vars))
  )
  # model.matrix: produce numeric matrix; uses training-level factors aligned via levels we set earlier
  mm <- model.matrix(~ . - 1, data = combined_preds)  # no intercept, all predictors
  # split back
  ntr <- nrow(train_df)
  train_mat <- mm[1:ntr, , drop = FALSE]
  test_mat  <- mm[(ntr + 1):nrow(mm), , drop = FALSE]
  
  # scale numeric columns using training means/sds to make kNN distance meaningful
  # identify numeric columns by checking which columns correspond to original numeric preds
  # safe approach: scale all columns (mean/sd from train) — if sd=0, set sd=1
  col_means <- colMeans(train_mat, na.rm = TRUE)
  col_sds   <- apply(train_mat, 2, sd, na.rm = TRUE)
  col_sds[col_sds == 0 | is.na(col_sds)] <- 1
  train_mat_scaled <- sweep(sweep(train_mat, 2, col_means, "-"), 2, col_sds, "/")
  test_mat_scaled  <- sweep(sweep(test_mat, 2, col_means, "-"), 2, col_sds, "/")
  
  
  if (nrow(train_mat_scaled) < 1 || nrow(test_mat_scaled) < 1) {
    warning("kNN: not enough data after filtering; returning NA for kNN")
    pred_knn <- rep(NA, length(true_values))
    acc_knn <- NA
    conf_knn <- NULL
  } else {
    k_used <- min(k_knn, nrow(train_mat_scaled))
    pred_knn_raw <- class::knn(train = train_mat_scaled, test = test_mat_scaled, cl = train_df[[target_var]], k = k_used)
    # place predictions back into full masked size (some rows may have been dropped if non-finite)
    pred_knn <- factor(pred_knn_raw, levels = target_levels)
    acc_knn <- mean(pred_knn == true_values, na.rm = TRUE)
    conf_knn <- table(Predicted = pred_knn, True = true_values)
  }
  
  # ---- 3) Random forest ----
  # Train randomForest on train_df, predict on test_df
  rf_model <- tryCatch({
    randomForest::randomForest(as.formula(paste(target_var, "~ .")), data = train_df, ntree = rf_ntree)
  }, error = function(e) {
    warning("randomForest failed: ", e$message)
    NULL
  })
  if (is.null(rf_model)) {
    pred_rf <- rep(NA, nrow(test_df))
    acc_rf <- NA
    conf_rf <- NULL
  } else {
    pred_rf_raw <- predict(rf_model, newdata = test_df)
    pred_rf <- factor(pred_rf_raw, levels = target_levels)
    acc_rf <- mean(pred_rf == true_values, na.rm = TRUE)
    conf_rf <- table(Predicted = pred_rf, True = true_values)
  }
  
  compute_precision_recall <- function(true, pred) {
  # Ensure both are factors with same levels
  true <- factor(true)
  pred <- factor(pred, levels = levels(true))
  
  cm <- caret::confusionMatrix(pred, true)
  precision <- mean(cm$byClass[, "Precision"], na.rm = TRUE)
  recall <- mean(cm$byClass[, "Recall"], na.rm = TRUE)
  
  return(list(precision = precision, recall = recall))
  }
  
  pr_mode <- compute_precision_recall(true_values, pred_mode)
  f1_mode <- f1(true_values, pred_mode)
  pr_knn  <- compute_precision_recall(true_values, pred_knn)
  f1_knn <- f1(true_values, pred_knn)
  pr_rf   <- compute_precision_recall(true_values, pred_rf)
  f1_rf <- f1(true_values, pred_rf)
  
  # Compose summary results
  summary_tbl <- tibble::tibble(
    Method = c("Mode", paste0("kNN (k=", k_knn, ")"), "RandomForest"),
    Accuracy = c(acc_mode, acc_knn, acc_rf),
    Precision = c(pr_mode$precision, pr_knn$precision, pr_rf$precision),
    Recall = c(pr_mode$recall, pr_knn$recall, pr_rf$recall),
    F1_score = c(f1_mode, f1_knn, f1_rf),
    N_test = nrow(test_df)
  )
  
  # Return details
  return(list(
    summary = summary_tbl,
    confusion = list(mode = conf_mode, knn = conf_knn, rf = conf_rf),
    true_values = true_values,
    preds = tibble(
      row_in_original = mask_idx,
      true = true_values,
      pred_mode = pred_mode,
      pred_knn = pred_knn,
      pred_rf = pred_rf
    ),
    models = list(rf_model = rf_model),
    train_info = list(n_train = nrow(train_df), predictors = predictor_vars)
  ))
}

result_BsmtFinType2 <- evaluate_categorical_imputation(train_csv, target_var = "BsmtFinType2",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123)
result_Electrical <- evaluate_categorical_imputation(train_csv, target_var = "Electrical",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123)
result_BsmtExposure <- evaluate_categorical_imputation(train_csv, target_var = "BsmtExposure",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123) 
print(result_BsmtFinType2$summary)
print(result_Electrical$summary)
print(result_BsmtExposure$summary)
```
Results of the imputation methods show that RandomForest consistently produces the most accurate results, although the differences in accuracy between methods aren't particularly large, which should be expected given the group sizes in these variables. In assessing the actual utility of these methods in assigning correctly, we should instead rely on other methods, such as the F1 score, or the harmonic mean of the models precision and recall. The F1 score also shows that RandomForest consistently produces the best results in terms of correctly assigning True positives / True negative cases, for which the mode and kNN are not as well suited. Knowing this, we'll now use the Random Forest model to impute the missing values from our original training dataset. The below function separates data into known / NA for train / test, and uses the given model (in this case Random Forests, although it could be changed for other methods) to impute the missing values in our original dataset.

```{r}
impute_categorical_var <- function(df, target_var, method = "rf", k_knn = 5, rf_ntree = 200) {
  
  # Identify missing and known rows
  missing_idx <- which(is.na(df[[target_var]]))
  known_idx   <- which(!is.na(df[[target_var]]))
  if (length(missing_idx) == 0) return(list(df = df, imputed = 0))
  if (length(known_idx) < 10)  return(list(df = df, imputed = 0)) # too few for model training
  
  # Select predictors 
  results_tbl <- feature_usefulness(train_csv, target_var = target_var)
  significant_predictors <- results_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
  predictors <- significant_predictors$Variable
  
  # Filter predictors to remove any with missing values
  complete_predictors <- predictors[sapply(df[predictors], function(x) all(!is.na(x)))]
  
  df_train <- df[known_idx, complete_predictors, drop = FALSE]
  df_test  <- df[missing_idx, complete_predictors, drop = FALSE]
  y_train  <- df[known_idx, target_var, drop = TRUE]
  
  # Model fitting and prediction
  preds <- NULL
  if (method == "mode") {
    mode_value <- names(sort(table(y_train), decreasing = TRUE))[1]
    preds <- factor(rep(mode_value, length(missing_idx)), levels = levels(y_train))
    
  } else if (method == "knn") {
    # Encode categoricals
    combined <- rbind(df_train, df_test)
    mm <- model.matrix(~ . - 1, data = combined)
    train_mat <- mm[1:nrow(df_train), , drop = FALSE]
    test_mat  <- mm[(nrow(df_train)+1):nrow(mm), , drop = FALSE]
    
    # Scale numeric columns
    col_means <- colMeans(train_mat)
    col_sds <- apply(train_mat, 2, sd)
    col_sds[col_sds == 0] <- 1
    train_mat <- sweep(sweep(train_mat, 2, col_means, "-"), 2, col_sds, "/")
    test_mat  <- sweep(sweep(test_mat, 2, col_means, "-"), 2, col_sds, "/")
    
    preds <- knn(train = train_mat, test = test_mat, cl = y_train, k = k_knn)
    
  } else if (method == "rf") {
    rf_model <- randomForest::randomForest(y = y_train, x = df_train, ntree = rf_ntree)
    preds <- predict(rf_model, newdata = df_test)
  } 
  
  # Replace missing values
  df[[target_var]][missing_idx] <- preds
  list(df = df, imputed = length(missing_idx))
}

impute_all_categorical <- function(df, method, k_knn = 5, rf_ntree = 200) {
  cat_vars <- names(df)[sapply(df, function(x) is.factor(x))]
  
  results <- list()
  total_imputed <- 0
  
  for (var in cat_vars) {
    if (any(is.na(df[[var]]))) {
      message("Imputing missing data in ", var, " using method = ", method, " ...")
      res <- impute_categorical_var(df, target_var = var, method = method,
                                    k_knn = k_knn, rf_ntree = rf_ntree)
      df <- res$df
      total_imputed <- total_imputed + res$imputed
      results[[var]] <- res$imputed
    }
  }
  
  message("Imputation complete. Total categorical values imputed: ", total_imputed)
  return(list(data = df, summary = results))
}
result <- impute_all_categorical(train_csv, method = "rf")
train_csv <- result$data
```

And remember to assign these as ordinal variables.

```{r}
train_csv$BsmtFinType2 <- factor(train_csv$BsmtFinType2, levels = BsmtFinType_levels, ordered = TRUE)
BsmtExposure_levels <- c('Missing', 'No', 'Mn', 'Av', 'Gd')
train_csv$BsmtExposure <- factor(train_csv$BsmtExposure, levels = BsmtExposure_levels, ordered = TRUE)
```

Now we're all done with the categorical variables, which brings us to our last variable, the continuous LotFrontage - defined as 'Linear feet of street connected to property'. We will first examine the shape of the variable

```{r}
train_density <- density(train_csv$LotFrontage, na.rm = TRUE)
plot(x= train_density$x, y =train_density$y, main= 'density of lot frontage', xlab = 'value', ylab = 'density')
boxplot(train_csv$LotFrontage, main= 'boxplot of lot frontage')
```
This data has a long right tail, there are outliers on both ends of the scale here, although there are a lot more large outliers rather than small outliers. This means that it might be difficult for imputing missing values by some measure of centrality in this dataset (i.e. mean or median) would likely lead to inaccuracies and problems further down the line.

Let's check for correlations between the lot frontage and other variables. There are several that I would expect that have some correlation, including lot shape, lot area, building type, and variables related to the size of the property, but specifically those related to the garage.

```{r}
num_cols <- names(train_csv)[sapply(train_csv, is.numeric)]
num_cols <- num_cols[-2] #removing LotFrontage
results_table_numeric <- tibble(
    Variable = character(0),
    correlation = 0,
    P_value = 0)
for (var in num_cols){
  corr <- cor.test(train_csv$LotFrontage, train_csv[[var]])
  results_table_numeric <- results_table_numeric %>%
        add_row(Variable = var, correlation = corr$estimate, P_value = corr$p.value)
}
results_table_numeric %>% arrange(P_value)
```
The top 10 variables that relate to lot frontage are the area of the first floor, the lot area, the living area, basement area, zoning subclass, garage area, total rooms above ground, sale price (not useful for us here as we'll also have to impute this in our test data where we won't have access to this variable), the number of cars that can fit in the garage, and the number of bedrooms above ground. This can broadly be interpreted as a larger house (i.e. one with more area in the total lot, more living area, more garage area, rooms above ground) can be correlated with an increase in lot frontage. Overall, there are 25 variables that reach a traditionally statistically significance level of association with a p-value < 0 .05.

We'll now try and evaluate the links between the other non-numerical variables and lot frontage. We'll conduct ANOVA or t-tests depending on the group sizes (most groups have more than 2 levels so will be using ANOVA) to evaluate whether the means of the groups differ by a significant amount.

```{r}
factor_vars <-names(train_csv)[sapply(train_csv, is.factor)]
results_table_anova <- tibble(
  Variable = character(0),
  F_value = 0,
  P_value = 0,
  Num_Groups = 0,
  eta_squared = 0
)
for (var in factor_vars) {
  n_levels <- nlevels(train_csv[[var]])
  data <- train_csv[[var]]
  if (length(unique(data)) > 2){
    formula <- as.formula(paste("LotFrontage ~", var))
    anova_results <- aov(formula, data = train_csv)
    eta_squared <- eta_squared(anova_results)
  #print(anova_results)
    results_table_anova <- results_table_anova %>%
      add_row(Variable = var, F_value = summary(anova_results)[1][[1]]$`F value`[1], P_value = summary(anova_results)[[1]]$`Pr(>F)`[1], Num_Groups = n_levels, eta_squared = eta_squared)
  }
  else{
    group_sizes <- table(train_csv[[var]])
    if (any(group_sizes==1)){
      cat(var, "skipped, not enough observations")
      next}
    else{
      formula <- as.formula(paste("LotFrontage ~", var))
      test <- t.test(formula, data = train_csv)
      eta_squared <- (test$statistic)^2 / ((test$statistic)^2 + (length(train_csv[[var]]) - 2))
      results_table_anova <- results_table_anova %>%
        add_row(Variable = var, F_value = test$statistic, P_value = test$p.value, Num_Groups = n_levels, eta_squared = eta_squared)
    }
  }
}

results_table_anova %>% arrange(P_value)

```
We get results of association for all variables except utilities, which has two levels with one level only having one observation. The results of the various ANOVA and t-tests show that the categorical variables broadly show some significant association with LotFrontage, 34 categorical variables reach a traditionally accepted p < 0.05 significantly associated threshold. Despite this, most variables show an eta_squared value (percentage of total variance accounted for by differences between the groups) that would generally be described as small, the top 11 show an eta > 0.06, which is traditionally accepted as a moderate effect. In any case, a good model will be able to take into account these smaller effects when imputing these values.

ANOVA's have several assumptions that are key to the results being interpretable, no significant outliers, equality of variances, and normal distribution, and these should be tested before we accept these results. If groups are normally distributed with equal variances, then a normal ANOVA would be reasonable, however if variances are unequal, we need to apply Welch's correction for an ANOVA, and if neither assumption is met then we need to use the Games-Howell test to assess for association. 

The below function gathers the dependent and outcome variable, removing empty observations, skipping groups with 2 or fewer groups, combines groups with small sizes (i.e. < 5 observations) into a new group ("Other"), rechecks to see if there are a sufficient number of groups (> 2) and skips variables with insufficient groups. The function then checks normality and equality of variances, and then carries out the appropriate test of association for the group.

```{r}
library(dplyr)
library(car)
library(stats)

check_anova_assumptions_adjusted <- function(df, outcome_var, cat_vars,
                                             alpha = 0.05,
                                             min_per_group = 3,
                                             min_obs_per_level = 5,
                                             make_plots = FALSE) {
  results <- list()
  
  for (var in cat_vars) {
    cat("\n--- Checking:", var, "---\n")
    tmp <- df %>% 
      select(all_of(c(outcome_var, var))) %>%
      na.omit()
    
    #Only check groups we need to, i.e. those with more than 2 groups
    if (nlevels(tmp[[var]]) < 3) {
      cat("Skipped:", var, "has < 3 levels.\n")
      next
    }
    if (any(table(tmp[[var]]) < min_per_group)) {
      cat("Warning:", var, "has groups with fewer than", min_per_group, "samples.\n")
    }
    
    
    valid_levels <- names(which(table(df[[var]]) >= min_obs_per_level))
    tmp[[var]] <- forcats::fct_other(tmp[[var]],
                                 keep = valid_levels,
                                 other_level = "Other")
    tmp[[var]] <- droplevels(tmp[[var]])
    
    level_counts2 <- table(tmp[[var]])
    valid_levels2 <- names(level_counts2[level_counts2 >= min_obs_per_level])
    tmp <- tmp[tmp[[var]] %in% valid_levels2, , drop = FALSE]
    tmp[[var]] <- droplevels(tmp[[var]])
    
    if (nlevels(tmp[[var]]) < 3) {
      cat("Skipped:", var, "has < 3 levels after filtering for small groups.\n")
      next
    }
    
    
    
    formula_str <- paste(outcome_var, "~", var)
    aov_model <- tryCatch(aov(as.formula(formula_str), data = tmp),
                          error = function(e) NULL)
    if (is.null(aov_model)) next
    
    resid_vals <- residuals(aov_model)
    
    # Assumption tests
    shapiro_p <- tryCatch(shapiro.test(resid_vals)$p.value, error = function(e) NA)
    lev_p <- tryCatch(leveneTest(as.formula(formula_str), data = tmp)[1, "Pr(>F)"],
                      error = function(e) NA)
    
    # Choose appropriate test
    normality_ok <- !is.na(shapiro_p) && shapiro_p > alpha
    equal_var_ok <- !is.na(lev_p) && lev_p > alpha
    
    recommended <- if (normality_ok && equal_var_ok) {
      "Standard ANOVA"
    } else if (normality_ok && !equal_var_ok) {
      "Welch’s ANOVA"
    } else {
      "Games–Howell"
    }
    
    # Run selected test
    p_value <- NA
    if (recommended == "Standard ANOVA") {
      p_value <- tryCatch(summary(aov_model)[[1]][["Pr(>F)"]][1], error = function(e) NA)
    } else if (recommended == "Welch’s ANOVA") {
      welch <- tryCatch(oneway.test(as.formula(formula_str), data = tmp, var.equal = FALSE),
                        error = function(e) NULL)
      p_value <- if (!is.null(welch)) welch$p.value else NA
    } else if (recommended == "Games–Howell" && nlevels(tmp[[var]]) > 2) {
      gh <- tryCatch(games_howell_test(formula = as.formula(formula_str), data = tmp),
                     error = function(e) NULL)
      p_value <- if (!is.null(gh)) gh$p.adj else NA
    }
    
    cat("Normality:", ifelse(normality_ok, "OK", "Violated"),
        "| Equal variances:", ifelse(equal_var_ok, "OK", "Violated"),
        "| → Using:", recommended, "| p =", round(p_value, 5), "\n")
    
    if (make_plots) {
      par(mfrow = c(1, 2))
      qqnorm(resid_vals, main = paste("QQ Plot for", var))
      qqline(resid_vals)
      boxplot(as.formula(formula_str), data = tmp, main = paste("Boxplot for", var),
              col = "lightblue", las = 2)
    }
    
    results[[var]] <- data.frame(
      Variable = var,
      Levels = nlevels(tmp[[var]]),
      Shapiro_p = shapiro_p,
      Levene_p = lev_p,
      Test_p = p_value,
      Normality_OK = normality_ok,
      EqualVar_OK = equal_var_ok,
      Recommended_Test = recommended,
      stringsAsFactors = FALSE
    )
    }
  
  results_df <- do.call(rbind, results)
  rownames(results_df) <- NULL
  
  fisher_df <- results_df %>%
    group_by(Variable) %>%
    summarise(
      n_pairs = n(),
      P_value= fisher(Test_p)$p,
      df = 2*n_pairs,
    )
  
  return(fisher_df %>% arrange(P_value))
}

cat_vars <- names(train_csv)[sapply(train_csv, is.factor)]
results_table_gh <- check_anova_assumptions_adjusted(
  df = train_csv,
  outcome_var = "LotFrontage",
  cat_vars = cat_vars,
  alpha = 0.05,
  make_plots = FALSE
)

print(results_table_gh)
```

All variables that did test used the Games-Howell test, indicating that neither the variables are normally distributed nor were the variances between groups equal. Two variables started with two levels (Street and CentralAir) so were not going to be included in the ANOVA regardless, and two more simplified down to a two sample test (MiscFeature and PoolQC) once very rare groups were removed. This test provides a more stringent and more statistically rigorous method of testing for association between the dependent variable (in this case LotFrontage) and the categorical variables which shows 18 variables meet a traditional p < 0.05 threshold compared to the 34 found when using a standard ANOVA on all categorical variables, although for the least strict threshold I'll use (0.1) there are a total of 23 variables found to have some association. 

Whilst this does produce a list of variables that we are more confident about claiming an association between the dependent and outcome variable, does this list of variables actually perform better than the list produced by ANOVA analysis when imputing the unknown values? We'll perform some testing now to check if this is the case. The next two functions will be the same, with one using ANOVA testing solely to check which categorical variables are more closely related to the outcome and one using Games-Howell testing to check for association. Beyond this, both functions aim to do the same thing, generate a list of variables that are associated with the outcome (LotFrontage in this case), then test different prediction methods, linear regression, random forest, and chained equations (MICE), with a combination of different parameter values for these methods as part of cross validation.

Given it's a long function, and it can take a few minutes to run (or longer if the parameter grid is expanded), I'll explain briefly what it specifically does, as this will form the basis of the imputation of the other continuous variables, the sale price. Firstly, the parameter grid is set by the user, by making the alpha values more stringent (i.e. 1e-12) we can restrict the association to the top ~30 most associated variables, compared to 59 at the least stringent standards of p < 0.05. The function aims to evaluate each possible combination of parameters possible so be careful when expanding the grid as each addition does make runtime increase significantly. These combinations are ran in parallel using the plan and future packages for this reason. Variables are selected by association as previously described, data is masked using a few safety methods to ensure completeness and validity of testing, and then prediction is carried out using the previously noted methods.

```{r}
set.seed(123)

# Use your dataframe name
df_all <- train_csv

# PARAM GRID (customise as needed)
alpha_cor_values    <- c(1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
alpha_anova_values  <- c(1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
missforest_ntree    <- c(50, 100, 150, 200)
mice_methods        <- c("cart")        # keep small list to limit runtime

# parallel plan: use most cores but leave one free
workers_to_use <- max(1, future::availableCores() - 1)
plan(multisession, workers = workers_to_use)

# make grid as a data.frame of combos
param_grid <- expand.grid(
  alpha_cor = alpha_cor_values,
  alpha_anova = alpha_anova_values,
  mf_ntree = missforest_ntree,
  mice_method = mice_methods,
  stringsAsFactors = FALSE
)

# The core evaluator
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 5,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 200) {

  # capture parameters
  alpha_cor   <- row_params$alpha_cor
  alpha_anova <- row_params$alpha_anova
  mf_ntree    <- row_params$mf_ntree
  mice_method <- row_params$mice_method
  
  # wrap in tryCatch so a single failing combo returns NA row
  tryCatch({
    df <- df  # use df_all by default
    
    # select numeric predictors by correlation 
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
   
    # select factor predictors by ANOVA 
    selected_factors <- results_table_anova$Variable[results_table_anova$P_value < alpha_anova]
    
    selected_predictors <- c(selected_numeric, selected_factors)
    if (length(selected_predictors) == 0) {
      # Return parameter settings
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_anova))
    }
    
    # keep only rows where predictors are complete (fair comparison) --
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]
    
    # known indices (where target known)
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      # not enough data to evaluate this combo properly -> return NA results
      return(tibble(
        alpha_cor = alpha_cor, alpha_anova = alpha_anova,
        en_alpha = en_alpha, mf_ntree = mf_ntree, mice_method = mice_method,
        Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = length(selected_predictors)
      ))
    }
    
    # mask creation (ensure training retains factor levels) --
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL
    
    # helper for mask validity
    # checks there are enough levels in the factors present 
    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]
    
    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      # fallback try smaller mask fractions
      found <- FALSE
      for (mf in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf))
        attempt2 <- 1
        while(attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
    }
    
    # Prepare masked dataframe 
    df_model_masked <- model_df
    true_values <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA  # mask
    
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final
    
    # Drop single-level factors or constant numerics in training
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) {
            final_predictors <- setdiff(final_predictors, f)
          }
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) {
          final_predictors <- setdiff(final_predictors, nvar)
        }
      }
    }
    
    # Build train/test frames
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]
    
    # Align factor levels: set test levels to training levels (prevents "new levels" error)
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)  # unseen => NA (safe)
      }
    }
    
    # -------- Linear regression --------
    lm_pred <- rep(NA_real_, length(test_idx))
    lm_success <- FALSE
    if (nrow(train_df_reg) >= 5 && length(final_predictors) > 0) {
      lm_formula <- as.formula(paste(target_var, "~", paste(final_predictors, collapse = " + ")))
      lm_fit <- tryCatch(lm(lm_formula, data = train_df_reg), error = function(e) e)
      if (!inherits(lm_fit, "error")) {
        # predict; rows with NA predictors will produce NA predictions
        lm_pred <- tryCatch(predict(lm_fit, newdata = test_df_reg), error = function(e) rep(NA_real_, nrow(test_df_reg)))
        lm_success <- TRUE
      }
    }
    
    # -------- MICE (masking already done in test_df_reg: LotFrontage is NA in test rows) --------
    mice_pred <- rep(NA_real_, length(test_idx))
    mice_success <- FALSE
    # Combine train+test (with test's LotFrontage currently NA) and impute
    combined_for_mice <- rbind(train_df_reg, test_df_reg)
    mice_result <- tryCatch(mice(combined_for_mice, m = 1, maxit = 5, method = mice_method, printFlag = FALSE),
                            error = function(e) NULL)
    if (!is.null(mice_result)) {
      mice_imputed <- complete(mice_result)
      ntr <- nrow(train_df_reg)
      mice_test  <- mice_imputed[(ntr + 1):nrow(mice_imputed), , drop = FALSE]
      mice_pred <- mice_test[[target_var]]
      mice_success <- TRUE
    }
    
    # -------- missForest (masking already done in test_df_reg) --------
    mf_pred <- rep(NA_real_, length(test_idx))
    mf_success <- FALSE
    combined_for_mf <- rbind(train_df_reg, test_df_reg)
    # ensure types are appropriate; missForest expects factors as factors and numerics numeric
    mf_out <- tryCatch(missForest(combined_for_mf, maxiter = 10, ntree = mf_ntree, verbose = TRUE),
                       error = function(e) NULL)
    if (!is.null(mf_out)) {
      mf_imputed <- mf_out$ximp
      ntr <- nrow(train_df_reg)
      mf_pred <- mf_imputed[[target_var]][(ntr + 1):nrow(mf_imputed)]
      mf_success <- TRUE
    }
    
    
    # Evaluate metrics using the true_values vector and each predicted vector
    eval_metrics <- function(true, pred) {
      ok <- !is.na(true) & !is.na(pred)
      if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
      t <- true[ok]; p <- pred[ok]
      list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
    }
    
    res_lm   <- eval_metrics(true_values, lm_pred)
    res_mice <- eval_metrics(true_values, mice_pred)
    res_mf   <- eval_metrics(true_values, mf_pred)
    
    # Compose results tibble for this combination
    result_tib <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "MICE", "missForest"),
      RMSE = c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE),
      MAE  = c(res_lm$MAE, res_mice$MAE, res_mf$MAE),
      R2   = c(res_lm$R2, res_mice$R2, res_mf$R2),
      n_predictors = length(final_predictors)
    )
    
    return(result_tib)
    
  }, error = function(e) {
    # On error, return NA result with the parameters
    tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_,
      n_predictors = NA_integer_
    )
  })
}

# Run grid in parallel (use reproducible random seeds)
# convert each row of param_grid to a list for mapping
param_list <- split(param_grid, seq_len(nrow(param_grid)))

# Use future_map_dfr with reproducible seeds
results_all_anova <- future_map_dfr(param_list, evaluate_combo, .options = furrr_options(seed = TRUE))

# Clean up: restore sequential plan if desired
plan(sequential)

# You can also inspect best combos:
best_by_method <- results_all_anova %>%
  group_by(Method) %>%
  filter(!is.na(RMSE)) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup()
print(best_by_method)
```
```{r}
set.seed(123)

# Use your dataframe name
df_all <- train_csv

# PARAM GRID (customise as needed)
alpha_cor_values    <- c(1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
alpha_gh_values     <- c(1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2, 1e-1)
missforest_ntree    <- c(50, 100, 150, 200, 250, 300, 350, 400)
mice_methods        <- c("cart")        # keep small list to limit runtime

# parallel plan: use most cores but leave one free
workers_to_use <- max(1, future::availableCores() - 1)
plan(multisession, workers = workers_to_use)

# make grid as a data.frame of combos
param_grid <- expand.grid(
  alpha_cor = alpha_cor_values,
  alpha_gh = alpha_gh_values,
  mf_ntree = missforest_ntree,
  mice_method = mice_methods,
  stringsAsFactors = FALSE
)

# The core evaluator
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 5,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 20) {
  # capture parameters
  alpha_cor   <- row_params$alpha_cor
  alpha_gh    <- row_params$alpha_gh
  mf_ntree    <- row_params$mf_ntree
  mice_method <- row_params$mice_method
  
  # wrap in tryCatch so a single failing combo returns NA row
  tryCatch({
    df <- df  # use df_all by default
    
    # select numeric predictors by correlation 
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
   
    # select factor predictors by Games-Howell 
    selected_factors <- results_table_gh$Variable[results_table_gh$P_value < alpha_gh]
    
    selected_predictors <- c(selected_numeric, selected_factors)
    if (length(selected_predictors) == 0) {
      # Return parameter settings
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_gh))
    }
    
    
      # keep only rows where predictors are complete 
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]
    
    # known indices (where target known)
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      # not enough data to evaluate this combo properly -> return NA results
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_gh))
    }
    
    # mask creation (ensure training retains factor levels) --
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL
    
    # helper for mask validity
    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]
    
    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      # fallback try smaller mask fractions
      found <- FALSE
      for (mf in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf))
        attempt2 <- 1
        while(attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
      if (!found) {
        return(sprintf('No masking found at masking fraction %d and max mask attempts %f, please adjust these values and try again', mask_frac, max_mask_attempts))
      }
    }
    
    # Prepare masked dataframe (same approach as your working code)
    df_model_masked <- model_df
    true_values <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA  # mask
    
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final
    
    # Drop single-level factors or constant numerics in training
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) {
            final_predictors <- setdiff(final_predictors, f)
          }
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) {
          final_predictors <- setdiff(final_predictors, nvar)
        }
      }
    }
    if (length(final_predictors) == 0) {
      return(tibble(
        alpha_cor = alpha_cor, alpha_gh = alpha_gh,
        en_alpha = en_alpha, mf_ntree = mf_ntree, mice_method = mice_method,
        Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = 0
      ))
    }
    
    # Build train/test frames
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]
    
    # Align factor levels: set test levels to training levels (prevents "new levels" error)
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)  # unseen => NA (safe)
      }
    }
    
    # -------- Linear regression --------
    lm_pred <- rep(NA_real_, length(test_idx))
    lm_success <- FALSE
    if (nrow(train_df_reg) >= 5 && length(final_predictors) > 0) {
      lm_formula <- as.formula(paste(target_var, "~", paste(final_predictors, collapse = " + ")))
      lm_fit <- tryCatch(lm(lm_formula, data = train_df_reg), error = function(e) e)
      if (!inherits(lm_fit, "error")) {
        # predict; rows with NA predictors will produce NA predictions
        lm_pred <- tryCatch(predict(lm_fit, newdata = test_df_reg), error = function(e) rep(NA_real_, nrow(test_df_reg)))
        lm_success <- TRUE
      }
    }
    
    # -------- MICE (masking already done in test_df_reg: LotFrontage is NA in test rows) --------
    mice_pred <- rep(NA_real_, length(test_idx))
    mice_success <- FALSE
    # Combine train+test (with test's LotFrontage currently NA) and impute
    combined_for_mice <- rbind(train_df_reg, test_df_reg)
    mice_result <- tryCatch(mice(combined_for_mice, m = 1, maxit = 5, method = mice_method, printFlag = FALSE),
                            error = function(e) NULL)
    if (!is.null(mice_result)) {
      mice_imputed <- complete(mice_result)
      ntr <- nrow(train_df_reg)
      mice_test  <- mice_imputed[(ntr + 1):nrow(mice_imputed), , drop = FALSE]
      mice_pred <- mice_test[[target_var]]
      mice_success <- TRUE
    }
    
    # -------- missForest (masking already done in test_df_reg) --------
    mf_pred <- rep(NA_real_, length(test_idx))
    mf_success <- FALSE
    combined_for_mf <- rbind(train_df_reg, test_df_reg)
    # ensure types are appropriate; missForest expects factors as factors and numerics numeric
    mf_out <- tryCatch(missForest(combined_for_mf, maxiter = 10, ntree = mf_ntree, verbose = FALSE),
                       error = function(e) NULL)
    if (!is.null(mf_out)) {
      mf_imputed <- mf_out$ximp
      ntr <- nrow(train_df_reg)
      mf_pred <- mf_imputed[[target_var]][(ntr + 1):nrow(mf_imputed)]
      mf_success <- TRUE
    }
    
    # Evaluate metrics using the true_values vector and each predicted vector
    eval_metrics <- function(true, pred) {
      ok <- !is.na(true) & !is.na(pred)
      if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
      t <- true[ok]; p <- pred[ok]
      list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
    }
    
    res_lm   <- eval_metrics(true_values, lm_pred)
    res_mice <- eval_metrics(true_values, mice_pred)
    res_mf   <- eval_metrics(true_values, mf_pred)
    
    # Compose results tibble for this combination
    result_tib <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_gh = alpha_gh,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "MICE", "missForest"),
      RMSE = c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE),
      MAE  = c(res_lm$MAE, res_mice$MAE, res_mf$MAE),
      R2   = c(res_lm$R2, res_mice$R2, res_mf$R2),
      n_predictors = length(final_predictors)
    )
    
    return(result_tib)
    
  }, error = function(e) {
    # On error, return NA result with the parameters
    tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_gh = alpha_gh,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_,
      n_predictors = NA_integer_
    )
  })
}

# Run grid in parallel (use reproducible random seeds)
# convert each row of param_grid to a list for mapping
param_list <- split(param_grid, seq_len(nrow(param_grid)))

# Use future_map_dfr with reproducible seeds
results_all_gh <- future_map_dfr(param_list, evaluate_combo, .options = furrr_options(seed = TRUE))

# Clean up: restore sequential plan if desired
plan(sequential)

# You can also inspect best combos:
best_by_method <- results_all_gh %>%
  group_by(Method) %>%
  filter(!is.na(RMSE)) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup()
print(best_by_method)
```
Again we see that the best predictions are found by combining as many predictors as we can (in this case 25 predictors) and using a random forest prediction model, achieving a RMSE of 11.74, MAE of 8.4, and R2 of 0.63. This is not much worse than the results found by using ANOVA alone to generate categorical variables associated with the outcome variable, but it is worse than the results found using ANOVA alone to associate categorical variables (RMSE = 10.36, R2 = 0.71). 

Deep learning models generally don't perform better than random forests on tabular imputation datasets like this, but just to check I've included one here. More for my sanity rather than any statistical completeness.

```{r}
# Load packages
library(dplyr)
library(forcats)
library(tidyr)
library(h2o)
library(rstatix)
library(tibble)

# Initialize H2O once
h2o.init(nthreads = 1, max_mem_size = "8G")

# ---- CONFIG ----
set.seed(123)
df_all <- train_csv
target_var <- "LotFrontage"

alpha_cor <- 0.01
alpha_gh  <- 0.01
mask_frac <- 0.2

# ---- Function ----
evaluate_h2o_deep <- function(df,
                              target_var = "LotFrontage",
                              alpha_cor = 0.01,
                              alpha_gh = 0.01,
                              mask_frac = 0.2,
                              min_obs_cor = 10,
                              min_obs_per_level = 5,
                              min_obs_anova = 20) {
  
  # --- Step 0: Clean and convert types ---
  df <- df %>%
    mutate(across(where(is.ordered), ~ factor(.x, ordered = FALSE))) %>%  # fix ordered factors
    mutate(across(where(is.character), as.factor))                        # ensure characters are factors
  
  # Identify variable types
  numeric_vars <- names(df)[sapply(df, is.numeric) & names(df) != target_var]
  factor_vars  <- names(df)[sapply(df, is.factor)]

  # --- Step 1: Select numeric predictors by correlation ---
  selected_numeric <- c()
  for (v in numeric_vars) {
    idx <- which(!is.na(df[[target_var]]) & !is.na(df[[v]]))
    if (length(idx) < min_obs_cor) next
    ct <- tryCatch(cor.test(df[[target_var]][idx], df[[v]][idx]), error = function(e) NULL)
    if (!is.null(ct) && !is.na(ct$p.value) && ct$p.value < alpha_cor) {
      selected_numeric <- c(selected_numeric, v)
    }
  }

  # --- Step 2: Select categorical predictors by Games-Howell test ---
  selected_factors <- c()
  for (v in factor_vars) {
    tmp <- df %>% select(all_of(c(target_var, v))) %>% drop_na()
    if (nrow(tmp) < min_obs_anova) next

    # Collapse rare levels
    valid_levels <- names(which(table(tmp[[v]]) >= min_obs_per_level))
    tmp[[v]] <- fct_other(tmp[[v]], keep = valid_levels, other_level = "Other")
    tmp[[v]] <- droplevels(tmp[[v]])

    if (nlevels(tmp[[v]]) < 2) next

    gh <- tryCatch(games_howell_test(as.formula(paste(target_var, "~", v)), data = tmp),
                   error = function(e) NULL)
    if (is.null(gh)) next

    p_values <- gh$p.adj
    p_values <- p_values[!is.na(p_values)]
    if (length(p_values) == 0) next
    chi_sq <- -2 * sum(log(p_values))
    combined_p <- 1 - pchisq(chi_sq, df = 2 * length(p_values))
    if (combined_p < alpha_gh) selected_factors <- c(selected_factors, v)
  }

  predictors <- c(selected_numeric, selected_factors)
  if (length(predictors) == 0) stop("No predictors selected; adjust alpha thresholds.")

  # --- Step 3: Prepare modeling data ---
  model_df <- df %>% select(all_of(c(target_var, predictors))) %>% drop_na()

  # Split known values
  known_idx <- which(!is.na(model_df[[target_var]]))
  mask_n <- max(1, round(length(known_idx) * mask_frac))
  mask_idx <- sample(known_idx, mask_n)

  train_df <- model_df[-mask_idx, ]
  test_df  <- model_df[mask_idx, ]

  true_values <- test_df[[target_var]]

  # --- Step 4: Convert to H2O frames ---
  train_h2o <- as.h2o(train_df)
  test_h2o  <- as.h2o(test_df)

  predictors <- setdiff(names(train_df), target_var)

  # --- Step 5: Deep Learning model ---
  grid_id <- paste0("dl_grid_", sample(1000:9999, 1))
  hyper_params <- list(
    hidden = list(c(32, 16), c(64, 32, 16), c(128, 64)),
    activation = c("Rectifier", "Tanh", "Maxout"),
    input_dropout_ratio = c(0, 0.1, 0.2),
    rate = c(0.01, 0.005, 0.001)
  )

  dl_grid <- h2o.grid(
    algorithm = "deeplearning",
    grid_id = grid_id,
    x = predictors,
    y = target_var,
    training_frame = train_h2o,
    validation_frame = test_h2o,
    hyper_params = hyper_params,
    epochs = 100,
    stopping_metric = "RMSE",
    stopping_rounds = 5,
    stopping_tolerance = 1e-4,
    seed = 123,
    standardize = TRUE
  )

  # --- Step 6: Best model and performance ---
  grid_perf <- h2o.getGrid(grid_id, sort_by = "RMSE", decreasing = FALSE)
  best_model_id <- grid_perf@model_ids[[1]]
  best_dl <- h2o.getModel(best_model_id)

  preds <- as.data.frame(h2o.predict(best_dl, test_h2o))[, 1]

  ok <- !is.na(true_values) & !is.na(preds)
  RMSE <- sqrt(mean((true_values[ok] - preds[ok])^2))
  MAE  <- mean(abs(true_values[ok] - preds[ok]))
  R2   <- cor(true_values[ok], preds[ok])^2

  tibble(
    Method = "H2O Deep Learning",
    alpha_cor = alpha_cor,
    alpha_gh = alpha_gh,
    RMSE = RMSE,
    MAE = MAE,
    R2 = R2,
    n_predictors = length(predictors)
  )
}

# ---- RUN ----
results_h2o <- evaluate_h2o_deep(df_all, alpha_cor = 0.01, alpha_gh = 0.01)
print(results_h2o)

# ---- CLEANUP ----
h2o.shutdown(prompt = FALSE)
```

Good for confirming my sanity, the deep learning model is performing worse on predicting this data when compared to a random forest model. 

Now we have our optimal prediction model and optimal parameters for this model, lets try to optimise it further. We'll start with increasing maxiter over a range, then attempt to find improvements via iterative refinement, and then evaluate individual features to see if removing them can provide improvements to the model. (note, tried this and none of it helped lmao)
```{r}
evaluate_ranger_imputation <- function(
    df,
    target_var = "LotFrontage",
    mask_frac = 0.2,
    min_train_rows = 30,
    ntree = 1000,
    mtry = NULL,
    min_node_size = 5,
    seed = 123
) {
  library(dplyr)
  library(ranger)
  
  set.seed(seed)

  # --- 1. Prepare data ---
  if (!target_var %in% names(df))
    stop("Target variable not found in dataframe.")
  
  # Ensure factors remain factors
  df[] <- lapply(df, function(x) if (is.character(x)) factor(x) else x)

  # Use only rows with non-missing target values
  known_idx <- which(!is.na(df[[target_var]]))
  if (length(known_idx) < min_train_rows + 2)
    stop("Not enough complete cases to perform evaluation.")

  # rows to mask
  mask_n <- max(1, round(mask_frac * length(known_idx)))
  mask_idx <- sample(known_idx, mask_n)

  # Keep true values
  true_vals <- df[[target_var]][mask_idx]

  # Create masked dataset
  df_masked <- df
  df_masked[[target_var]][mask_idx] <- NA
  
  # define train/test
  train_idx <- setdiff(known_idx, mask_idx)
  test_idx  <- mask_idx

  train_df <- df_masked[train_idx, , drop = FALSE]
  test_df  <- df_masked[test_idx, , drop = FALSE]

  # remove predictors that become all-NA or zero variance
  usable_predictors <- setdiff(names(df), target_var)

  usable_predictors <- usable_predictors[
    sapply(usable_predictors, function(v) {
      x <- train_df[[v]]
      if (all(is.na(x))) return(FALSE)
      if (is.numeric(x) && length(unique(na.omit(x))) <= 1) return(FALSE)
      if (is.factor(x) && nlevels(droplevels(x)) <= 1) return(FALSE)
      TRUE
    })
  ]

  if (length(usable_predictors) == 0)
    stop("No usable predictors remain after cleaning.")
  usable_predictors <- intersect(usable_predictors, final_predictors)

  # Keep only final predictors
  train_df <- train_df[, c(target_var, usable_predictors), drop = FALSE]
  test_df  <- test_df[, c(target_var, usable_predictors), drop = FALSE]

  # Align factor levels in test set
  for (col in usable_predictors) {
    if (is.factor(train_df[[col]])) {
      levs <- levels(droplevels(train_df[[col]]))
      train_df[[col]] <- factor(train_df[[col]], levels = levs)
      test_df[[col]]  <- factor(test_df[[col]],  levels = levs)
    }
  }

  # --- 2. Fit ranger model ---
  formula <- as.formula(paste(target_var, "~ ."))

  ranger_fit <- ranger(
    formula,
    data = train_df,
    num.trees = ntree,
    mtry = if (is.null(mtry)) floor(sqrt(length(usable_predictors))) else mtry,
    min.node.size = min_node_size,
    importance = "impurity",
    seed = seed
  )

  # --- 3. Predictions ---
  pred <- predict(ranger_fit, data = test_df)$predictions

  # --- 4. Metrics ---
  ok <- !is.na(true_vals) & !is.na(pred)
  if (sum(ok) == 0) {
    return(list(
      RMSE = NA, MAE = NA, R2 = NA,
      n_predictors = length(usable_predictors),
      n_test = length(true_vals)
    ))
  }

  y <- true_vals[ok]
  p <- pred[ok]

  RMSE <- sqrt(mean((y - p)^2))
  MAE  <- mean(abs(y - p))
  R2   <- if (length(y) > 1) cor(y, p)^2 else NA

  return(list(
    RMSE = RMSE,
    MAE = MAE,
    R2 = R2,
    n_predictors = length(usable_predictors),
    n_test = length(y),
    predictors = usable_predictors
  ))
}

result <- evaluate_ranger_imputation(train_csv, target_var = "LotFrontage")

result
```
```

```{r}
na_count_test
```
Missing data is present in the MSZoning, LotFrontage, Alley, Utilities, Exterior1st, Exterior2nd, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinSF1, BsmtFinType2, BsmtFinSF2, BsmtUnfSF, TotalBsmtSF, BsmtFullBath, BsmtHalfBath, KitchenQual, Functional, FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageCars, GarageArea, GarageQual, GarageCond, PoolQC, Fence, MiscFeature, SaleType
For the sake of consistency, we will use broadly the same methods as used in the training dataset to impute the missing values in the test dataset.