---
title: "data exploration for house prices dataset"
output: html_document
date: "2025-10-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the required packages for this exploration. The code should run and
install packages if they are not installed on the local device

```{r}
requiredPackages = c('here', 'tidyverse', 'dplyr', 'data.table', 'ggplot2', 'rstatix', 'missForest', 'Metrics', 'glmnet', 'xgboost', 'mice', 'furrr', 'ufs', 'forcats', 'randomForest', 'future.apply', 'recipes', 'caret', 'broom', 'DescTools', 'infotheo', 'tibble', 'car', 'stats', 'poolr', 'keras', 'tensorflow', 'yardstick', 'h2o', 'ranger', 'lightgbm', 'Matrix', 'catboost', 'mgcv')
for (p in requiredPackages){
  if(!require(p,character.only = TRUE)) install.packages(p)
  library(p, character.only = TRUE)}
```

Here we're loading in the data required for this project

```{r}
setwd('C:/Users/marnt/Desktop/kaggle/house prices')
here() #Remember to reset the working directory as appropriate
train_csv <-as.data.frame(read.csv('train.csv'))
test_csv <- as.data.frame(read.csv('test.csv'))

```

We'll need to check the quality of the data before we can start predicting the sale prices. First check is to see where data is missing

```{r}
na_count_train <-sapply(train_csv, function(y) sum(length(which(is.na(y)))))
na_count_test <-sapply(test_csv, function(y) sum(length(which(is.na(y)))))
```

``` {r}
na_count_train
```
Missing data present in LotFrontage, Alley, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, FireplaceQu, GarageType, GarageYrBlt, GarageQual, GarageCond, PoolQC, Fence and MiscFeature fields.
We will need to further examine the data to decide whether or not to impute these missing values in some way

As a starting point, with missing values in the majority of entries for Alley, PoolQC, Fence, MiscFeature and FireplaceQu, we'll first investigate these features to determine whether they should be removed.


```{r}
train_csv %>% count(Alley)
train_csv %>% count(PoolQC)
train_csv %>% count(Fence)
train_csv %>% count(MiscFeature)
train_csv %>% count(FireplaceQu)
```
Missing values in FireplaceQu correspond to the property not having a fireplace, and we can impute missing values here easily. We will insert a new category to show that the properties do not have a fireplace. Imputing the missing values in other fields is similar in that we are assuming a missing value indicates the feature not being present on the property. This system applies similarly to the missing values in the fields about basements, masonry, and garages, although we will use a slightly different approach for those fields.

Note that there are two entries here in the basement field where information is missing for in just one basement field (BsmtExposure and BsmtFinType2) and not the other basement fields, which will be imputed by other methods.

```{r}
missing_cols_categorical <- c('FireplaceQu', 'Alley', 'PoolQC', 'Fence', 'MiscFeature', 'MasVnrType', 'GarageQual', 'GarageFinish', 'GarageCond', 'GarageType')
missing_cols_numerical <- c('MasVnrArea', 'GarageYrBlt')
for (col in missing_cols_categorical){
  train_csv[[col]][is.na(train_csv[[col]])] <- 'Missing'
}
for (col in missing_cols_numerical){
  train_csv[[col]][is.na(train_csv[[col]])] <- 0
}

bsmt_cols <- c('BsmtQual', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2')
all_missing <- apply(train_csv[, bsmt_cols], 1, function(row) all(is.na(row)))
train_csv[all_missing, bsmt_cols] <- 'Missing'
```
Before we start imputing values for variables which are missing data, like LotFrontage, and those associated with other house features, we'll need to further examine other features in the data.

Some variables are listed as numerical, but are in fact categorical and should be treated as such. These include the month that the property was sold (MoSold), the year the property was sold (YrSold) and the Subclass of the building (MSSubClass).

```{r}
train_csv$MoSold = as.factor(train_csv$MoSold)
train_csv$YrSold = as.factor(train_csv$YrSold)
train_csv$MSSubClass = as.factor(train_csv$MSSubClass)
```

This brings up another problem, some of the categorical variables in the dataset are ordinal and some are nominal, and they should be encoded as such to avoid problems further down the imputation pipeline. Some features (BsmtFinType2, BsmtExposure) have missing data that will be imputed so we will classify them as ordered after the imputation.

```{r}
#First, make everything that's text into a factor
character_cols <- lapply(train_csv, class) == 'character'
train_csv[,character_cols] <- lapply(train_csv[,character_cols], as.factor)

#Next, make sure that everything that has a scale (i.e. factors that are ordinal) has that scale defined in order
factor_vars <- lapply(train_csv, class) == 'factor'

#These are the ordinal factors that use the same scale, the other 7 need to be set individually
ordinal_vars <- c('ExterQual', 'ExterCond', 'BsmtQual', 'BsmtCond', 'HeatingQC', 'KitchenQual', 'FireplaceQu', 'GarageQual', 'GarageCond', 'PoolQC')                      
quality_levels <- c('Missing', 'Po', 'Fa', 'TA', 'Gd', 'Ex')
for (var in ordinal_vars){
  train_csv[[var]] <- factor(train_csv[[var]], levels = quality_levels, ordered = TRUE) 
}

#These are the other ordinal variables with their unique scales
BsmtFinType_levels <- c('Missing', 'Unf', 'LwQ', 'Rec', 'BLQ', 'ALQ', 'GLQ')
train_csv$BsmtFinType1 <- factor(train_csv$BsmtFinType1, levels = BsmtFinType_levels, ordered = TRUE)
Functional_levels <- c('Sal','Sev', 'Maj2', 'Maj1', 'Mod', 'Min2', 'Min1', 'Typ')
train_csv$Functional <- factor(train_csv$Functional, levels = Functional_levels, ordered = TRUE)
GarageFinish_levels <- c('Missing', 'Unf', 'RFn', 'Fin')
train_csv$GarageFinish <- factor(train_csv$GarageFinish, levels = GarageFinish_levels, ordered = TRUE)
Fence_levels <- c('Missing', 'MnWw', 'GdWo', 'MnPrv', 'GdPrv')
train_csv$Fence <- factor(train_csv$Fence, levels = Fence_levels, ordered = TRUE)
Slope_levels <- c('Sev', 'Mod', 'Gtl')
train_csv$LandSlope <- factor(train_csv$LandSlope, levels = Slope_levels, ordered = TRUE)
Paved_levels <- c('N', 'P', 'Y')
train_csv$PavedDrive <- factor(train_csv$PavedDrive, levels = Paved_levels, ordered = TRUE)
```

Before we go ahead and predict our missing values, we'll first try and examine the variables in the dataset and see which variables are strongly associated or un-associated with our outcome variables. We'll use Cramers V test for categorical variables and ANOVA tests for numerical variables.


```{r}
feature_usefulness <- function(df, target_var) {
  target <- df[[target_var]]
  
  # Drop NAs in target for association calculation
  df <- df[!is.na(target), ]
  target <- droplevels(df[[target_var]])
  
  results <- data.frame(
    Variable = character(),
    Type = character(),
    Measure = character(),
    Value = numeric(),
    stringsAsFactors = FALSE
  )
  
  for (v in setdiff(names(df), target_var)) {
    # Skip variables that are entirely missing or constant
    if (all(is.na(df[[v]])) || length(unique(df[[v]])) < 2) next
    
    if (is.factor(df[[v]])) {
      # Categorical predictor — use Cramér’s V
      tbl <- table(df[[v]], target)
      val <- tryCatch(DescTools::CramerV(tbl, bias.correct = TRUE), error = function(e) NA)
      results <- rbind(results, data.frame(Variable = v, Type = "Categorical", Measure = "CramerV", Value = val))
      
    } else if (is.numeric(df[[v]])) {
      # Numeric predictor — use correlation ratio (η²)
      formula <- as.formula(paste(v, "~", target_var))
      val <- tryCatch({
        model <- aov(formula, data = df)
        summary(model)[[1]][["Sum Sq"]][1] / sum(summary(model)[[1]][["Sum Sq"]])
      }, error = function(e) NA)
      results <- rbind(results, data.frame(Variable = v, Type = "Numeric", Measure = "EtaSquared", Value = val))
    }
  }
  
  results <- results %>%
    arrange(desc(Value))
  
  return(results)
}

usefulness_tbl <- feature_usefulness(train_csv, target_var = "BsmtExposure")
significant_predictors <- usefulness_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
```

Now that we've got most of the missing categorical data appropriately assigned and imputed, we need to deal with the remaining missing values. Accomplishing this task will involve testing different methods to see which method provides the best results for imputation, as assessed by different methods (Accuracy of predictions, F1 Score). This function does several things that I'll describe briefly. First, there are a few safety checks that are completed as we select our significantly associated predictors (although this step is really more about removing very unassociated predictors), before a masked set of data is taken from the dataset and used to split the data into training / testing. We impute the values missing from the categories not being tested for (at this point LotFrontage, and the other categorical variables that are being imputed).Then with our cleaned data, we predict! As a baseline, we're using the mode values from the category as the outcome, which given the group sizes in these categories isn't a very bad idea. For two comparison methods, we'll use kNN and Random Forest classification models.

```{r}
evaluate_categorical_imputation <- function(df,
                                            target_var,
                                            predictor_vars = NULL,
                                            mask_frac = 0.20,
                                            k_knn = 5,
                                            rf_ntree = 200,
                                            seed = 123,
                                            max_predictor_missing_frac = 0.5) {
  set.seed(seed)
  if (!target_var %in% names(df)) stop("target_var not found in dataframe")
  # ensure target is factor
  if (!is.factor(df[[target_var]])) df[[target_var]] <- as.factor(df[[target_var]])
  
  results_tbl <- feature_usefulness(train_csv, target_var = target_var)
  significant_predictors <- results_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
  
  
  # choose predictors if not provided: all columns except target
  if (is.null(predictor_vars)) {
    predictor_vars <- significant_predictors$Variable
  }
  
  # Filter out predictors that are almost entirely missing
  predictor_missing_frac <- sapply(df[predictor_vars], function(x) mean(is.na(x)))
  keep_preds <- names(predictor_missing_frac)[predictor_missing_frac <= max_predictor_missing_frac]
  if (length(keep_preds) == 0) stop("No predictors with acceptable missingness found")
  predictor_vars <- keep_preds
  
  # Split known / missing target rows
  known_idx <- which(!is.na(df[[target_var]]))
  missing_target_idx <- which(is.na(df[[target_var]]))
  if (length(known_idx) < 10) stop("Too few known target rows to evaluate reliably")
  
  # Mask a random subset of known targets for evaluation
  mask_n <- max(1, round(length(known_idx) * mask_frac))
  mask_idx <- sample(known_idx, mask_n)
  df_masked <- df
  true_values <- df[[target_var]][mask_idx]
  df_masked[[target_var]][mask_idx] <- NA
  
  # Build training and testframes:
  # training rows = rows with target known (after masking)
  train_df <- df_masked %>% filter(!is.na(.data[[target_var]])) %>% select(all_of(c(target_var, predictor_vars)))
  test_df  <- df_masked[mask_idx, , drop = FALSE] %>% select(all_of(c(target_var, predictor_vars)))
  
  # Apply simple imputation to predictors using only training distribution for numerics:
  # For numeric medians we compute medians on training data and apply to both train/test,
  # For categorical we map NAs to "Missing".
  # Compute numeric medians from training.
  numeric_preds <- predictor_vars[sapply(train_df[predictor_vars], is.numeric)]
  medians <- sapply(train_df[numeric_preds], function(x) median(x, na.rm = TRUE))
  for (col in numeric_preds){
  train_df[[col]][is.na(train_df[[col]])] <- medians[[col]]
  test_df[[col]][is.na(test_df[[col]])] <- medians[[col]]
  }
  
  factor_preds <- setdiff(predictor_vars, numeric_preds)
  for (col in factor_preds){
    if(NA %in% train_df[[col]]){
      levs <- c(levels(train_df[[col]]), 'MissingVal')
      train_df[[col]] <- factor(train_df[[col]], levels = levs)
      test_df[[col]] <- factor(test_df[[col]], levels = levs)
    }
    train_df[[col]][is.na(train_df[[col]])] <- 'MissingVal'
  
    if(NA %in% test_df[[col]]){
      levs <- c(levels(test_df[[col]]), 'MissingVal')
      train_df[[col]] <- factor(train_df[[col]], levels = levs)
      test_df[[col]] <- factor(test_df[[col]], levels = levs)
    }
    test_df[[col]][is.na(test_df[[col]])] <- 'MissingVal'
  }
  
  # If medians are NA (all missing), set to 0
  medians[is.na(medians)] <- 0
  
  # Ensure target in train is factor and levels set
  train_df[[target_var]] <- factor(train_df[[target_var]])
  target_levels <- levels(train_df[[target_var]])
  
  # ---- 1) Baseline: Mode ----
  mode_value <- names(sort(table(train_df[[target_var]]), decreasing = TRUE))[1]
  pred_mode <- factor(rep(mode_value, nrow(test_df)), levels = target_levels)
  acc_mode <- mean(pred_mode == true_values, na.rm = TRUE)
  conf_mode <- table(Predicted = pred_mode, True = true_values)
  
  # ---- 2) kNN ----
  # Build dummy variables using model.matrix on combined train+test so columns match
  combined_preds <- bind_rows(
    train_df %>% select(all_of(predictor_vars)),
    test_df %>% select(all_of(predictor_vars))
  )
  # model.matrix: produce numeric matrix; uses training-level factors aligned via levels we set earlier
  mm <- model.matrix(~ . - 1, data = combined_preds)  # no intercept, all predictors
  # split back
  ntr <- nrow(train_df)
  train_mat <- mm[1:ntr, , drop = FALSE]
  test_mat  <- mm[(ntr + 1):nrow(mm), , drop = FALSE]
  
  # scale numeric columns using training means/sds to make kNN distance meaningful
  # identify numeric columns by checking which columns correspond to original numeric preds
  # safe approach: scale all columns (mean/sd from train) — if sd=0, set sd=1
  col_means <- colMeans(train_mat, na.rm = TRUE)
  col_sds   <- apply(train_mat, 2, sd, na.rm = TRUE)
  col_sds[col_sds == 0 | is.na(col_sds)] <- 1
  train_mat_scaled <- sweep(sweep(train_mat, 2, col_means, "-"), 2, col_sds, "/")
  test_mat_scaled  <- sweep(sweep(test_mat, 2, col_means, "-"), 2, col_sds, "/")
  
  
  if (nrow(train_mat_scaled) < 1 || nrow(test_mat_scaled) < 1) {
    warning("kNN: not enough data after filtering; returning NA for kNN")
    pred_knn <- rep(NA, length(true_values))
    acc_knn <- NA
    conf_knn <- NULL
  } else {
    k_used <- min(k_knn, nrow(train_mat_scaled))
    pred_knn_raw <- class::knn(train = train_mat_scaled, test = test_mat_scaled, cl = train_df[[target_var]], k = k_used)
    # place predictions back into full masked size (some rows may have been dropped if non-finite)
    pred_knn <- factor(pred_knn_raw, levels = target_levels)
    acc_knn <- mean(pred_knn == true_values, na.rm = TRUE)
    conf_knn <- table(Predicted = pred_knn, True = true_values)
  }
  
  # ---- 3) Random forest ----
  # Train randomForest on train_df, predict on test_df
  rf_model <- tryCatch({
    randomForest::randomForest(as.formula(paste(target_var, "~ .")), data = train_df, ntree = rf_ntree)
  }, error = function(e) {
    warning("randomForest failed: ", e$message)
    NULL
  })
  if (is.null(rf_model)) {
    pred_rf <- rep(NA, nrow(test_df))
    acc_rf <- NA
    conf_rf <- NULL
  } else {
    pred_rf_raw <- predict(rf_model, newdata = test_df)
    pred_rf <- factor(pred_rf_raw, levels = target_levels)
    acc_rf <- mean(pred_rf == true_values, na.rm = TRUE)
    conf_rf <- table(Predicted = pred_rf, True = true_values)
  }
  
  compute_precision_recall <- function(true, pred) {
  # Ensure both are factors with same levels
  true <- factor(true)
  pred <- factor(pred, levels = levels(true))
  
  cm <- caret::confusionMatrix(pred, true)
  precision <- mean(cm$byClass[, "Precision"], na.rm = TRUE)
  recall <- mean(cm$byClass[, "Recall"], na.rm = TRUE)
  
  return(list(precision = precision, recall = recall))
  }
  
  pr_mode <- compute_precision_recall(true_values, pred_mode)
  f1_mode <- f1(true_values, pred_mode)
  pr_knn  <- compute_precision_recall(true_values, pred_knn)
  f1_knn <- f1(true_values, pred_knn)
  pr_rf   <- compute_precision_recall(true_values, pred_rf)
  f1_rf <- f1(true_values, pred_rf)
  
  # Compose summary results
  summary_tbl <- tibble::tibble(
    Method = c("Mode", paste0("kNN (k=", k_knn, ")"), "RandomForest"),
    Accuracy = c(acc_mode, acc_knn, acc_rf),
    Precision = c(pr_mode$precision, pr_knn$precision, pr_rf$precision),
    Recall = c(pr_mode$recall, pr_knn$recall, pr_rf$recall),
    F1_score = c(f1_mode, f1_knn, f1_rf),
    N_test = nrow(test_df)
  )
  
  # Return details
  return(list(
    summary = summary_tbl,
    confusion = list(mode = conf_mode, knn = conf_knn, rf = conf_rf),
    true_values = true_values,
    preds = tibble(
      row_in_original = mask_idx,
      true = true_values,
      pred_mode = pred_mode,
      pred_knn = pred_knn,
      pred_rf = pred_rf
    ),
    models = list(rf_model = rf_model),
    train_info = list(n_train = nrow(train_df), predictors = predictor_vars)
  ))
}

result_BsmtFinType2 <- evaluate_categorical_imputation(train_csv, target_var = "BsmtFinType2",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123)
result_Electrical <- evaluate_categorical_imputation(train_csv, target_var = "Electrical",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123)
result_BsmtExposure <- evaluate_categorical_imputation(train_csv, target_var = "BsmtExposure",
                                           predictor_vars = NULL, mask_frac = 0.2,
                                           k_knn = 15, rf_ntree = 225, seed = 123) 
print(result_BsmtFinType2$summary)
print(result_Electrical$summary)
print(result_BsmtExposure$summary)
```
Results of the imputation methods show that RandomForest consistently produces the most accurate results, although the differences in accuracy between methods aren't particularly large, which should be expected given the group sizes in these variables. In assessing the actual utility of these methods in assigning correctly, we should instead rely on other methods, such as the F1 score, or the harmonic mean of the models precision and recall. The F1 score also shows that RandomForest consistently produces the best results in terms of correctly assigning True positives / True negative cases, for which the mode and kNN are not as well suited. Knowing this, we'll now use the Random Forest model to impute the missing values from our original training dataset. The below function separates data into known / NA for train / test, and uses the given model (in this case Random Forests, although it could be changed for other methods) to impute the missing values in our original dataset.

```{r}
impute_categorical_var <- function(df, target_var, method = "rf", k_knn = 5, rf_ntree = 200) {
  
  # Identify missing and known rows
  missing_idx <- which(is.na(df[[target_var]]))
  known_idx   <- which(!is.na(df[[target_var]]))
  if (length(missing_idx) == 0) return(list(df = df, imputed = 0))
  if (length(known_idx) < 10)  return(list(df = df, imputed = 0)) # too few for model training
  
  # Select predictors 
  results_tbl <- feature_usefulness(train_csv, target_var = target_var)
  significant_predictors <- results_tbl %>%
  filter((Measure == "CramerV" & Value > 0.1) |
         (Measure == "EtaSquared" & Value > 0.05))
  predictors <- significant_predictors$Variable
  
  # Filter predictors to remove any with missing values
  complete_predictors <- predictors[sapply(df[predictors], function(x) all(!is.na(x)))]
  
  df_train <- df[known_idx, complete_predictors, drop = FALSE]
  df_test  <- df[missing_idx, complete_predictors, drop = FALSE]
  y_train  <- df[known_idx, target_var, drop = TRUE]
  
  # Model fitting and prediction
  preds <- NULL
  if (method == "mode") {
    mode_value <- names(sort(table(y_train), decreasing = TRUE))[1]
    preds <- factor(rep(mode_value, length(missing_idx)), levels = levels(y_train))
    
  } else if (method == "knn") {
    # Encode categoricals
    combined <- rbind(df_train, df_test)
    mm <- model.matrix(~ . - 1, data = combined)
    train_mat <- mm[1:nrow(df_train), , drop = FALSE]
    test_mat  <- mm[(nrow(df_train)+1):nrow(mm), , drop = FALSE]
    
    # Scale numeric columns
    col_means <- colMeans(train_mat)
    col_sds <- apply(train_mat, 2, sd)
    col_sds[col_sds == 0] <- 1
    train_mat <- sweep(sweep(train_mat, 2, col_means, "-"), 2, col_sds, "/")
    test_mat  <- sweep(sweep(test_mat, 2, col_means, "-"), 2, col_sds, "/")
    
    preds <- knn(train = train_mat, test = test_mat, cl = y_train, k = k_knn)
    
  } else if (method == "rf") {
    rf_model <- randomForest::randomForest(y = y_train, x = df_train, ntree = rf_ntree)
    preds <- predict(rf_model, newdata = df_test)
  } 
  
  # Replace missing values
  df[[target_var]][missing_idx] <- preds
  list(df = df, imputed = length(missing_idx))
}

impute_all_categorical <- function(df, method, k_knn = 5, rf_ntree = 200) {
  cat_vars <- names(df)[sapply(df, function(x) is.factor(x))]
  
  results <- list()
  total_imputed <- 0
  
  for (var in cat_vars) {
    if (any(is.na(df[[var]]))) {
      message("Imputing missing data in ", var, " using method = ", method, " ...")
      res <- impute_categorical_var(df, target_var = var, method = method,
                                    k_knn = k_knn, rf_ntree = rf_ntree)
      df <- res$df
      total_imputed <- total_imputed + res$imputed
      results[[var]] <- res$imputed
    }
  }
  
  message("Imputation complete. Total categorical values imputed: ", total_imputed)
  return(list(data = df, summary = results))
}
result <- impute_all_categorical(train_csv, method = "rf")
train_csv <- result$data
```

And remember to assign these as ordinal variables.

```{r}
train_csv$BsmtFinType2 <- factor(train_csv$BsmtFinType2, levels = BsmtFinType_levels, ordered = TRUE)
BsmtExposure_levels <- c('Missing', 'No', 'Mn', 'Av', 'Gd')
train_csv$BsmtExposure <- factor(train_csv$BsmtExposure, levels = BsmtExposure_levels, ordered = TRUE)
```

Now we're all done with the categorical variables, which brings us to our last variable, the continuous LotFrontage - defined as 'Linear feet of street connected to property'. We will first examine the shape of the variable

```{r}
train_density <- density(train_csv$LotFrontage, na.rm = TRUE)
plot(x= train_density$x, y =train_density$y, main= 'density of lot frontage', xlab = 'value', ylab = 'density')
boxplot(train_csv$LotFrontage, main= 'boxplot of lot frontage')
```
This data has a long right tail, there are outliers on both ends of the scale here, although there are a lot more large outliers rather than small outliers. This means that it might be difficult for imputing missing values by some measure of centrality in this dataset (i.e. mean or median) would likely lead to inaccuracies and problems further down the line.

Let's check for correlations between the lot frontage and other variables. There are several that I would expect that have some correlation, including lot shape, lot area, building type, and variables related to the size of the property, but specifically those related to the garage.

```{r}
num_cols <- names(train_csv)[sapply(train_csv, is.numeric)]
num_cols <- num_cols[-2] #removing LotFrontage
results_table_numeric <- tibble(
    Variable = character(0),
    correlation = 0,
    P_value = 0)
for (var in num_cols){
  corr <- cor.test(train_csv$LotFrontage, train_csv[[var]])
  results_table_numeric <- results_table_numeric %>%
        add_row(Variable = var, correlation = corr$estimate, P_value = corr$p.value)
}
results_table_numeric %>% arrange(P_value)
```
The top 10 variables that relate to lot frontage are the area of the first floor, the lot area, the living area, basement area, zoning subclass, garage area, total rooms above ground, sale price (not useful for us here as we'll also have to impute this in our test data where we won't have access to this variable), the number of cars that can fit in the garage, and the number of bedrooms above ground. This can broadly be interpreted as a larger house (i.e. one with more area in the total lot, more living area, more garage area, rooms above ground) can be correlated with an increase in lot frontage. Overall, there are 25 variables that reach a traditionally statistically significance level of association with a p-value < 0 .05.

We'll now try and evaluate the links between the other non-numerical variables and lot frontage. We'll conduct ANOVA or t-tests depending on the group sizes (most groups have more than 2 levels so will be using ANOVA) to evaluate whether the means of the groups differ by a significant amount.

```{r}
factor_vars <-names(train_csv)[sapply(train_csv, is.factor)]
results_table_anova <- tibble(
  Variable = character(0),
  F_value = 0,
  P_value = 0,
  Num_Groups = 0,
  eta_squared = 0
)
for (var in factor_vars) {
  n_levels <- nlevels(train_csv[[var]])
  data <- train_csv[[var]]
  if (length(unique(data)) > 2){
    formula <- as.formula(paste("LotFrontage ~", var))
    anova_results <- aov(formula, data = train_csv)
    eta_squared <- eta_squared(anova_results)
  #print(anova_results)
    results_table_anova <- results_table_anova %>%
      add_row(Variable = var, F_value = summary(anova_results)[1][[1]]$`F value`[1], P_value = summary(anova_results)[[1]]$`Pr(>F)`[1], Num_Groups = n_levels, eta_squared = eta_squared)
  }
  else{
    group_sizes <- table(train_csv[[var]])
    if (any(group_sizes==1)){
      cat(var, "skipped, not enough observations")
      next}
    else{
      formula <- as.formula(paste("LotFrontage ~", var))
      test <- t.test(formula, data = train_csv)
      eta_squared <- (test$statistic)^2 / ((test$statistic)^2 + (length(train_csv[[var]]) - 2))
      results_table_anova <- results_table_anova %>%
        add_row(Variable = var, F_value = test$statistic, P_value = test$p.value, Num_Groups = n_levels, eta_squared = eta_squared)
    }
  }
}

results_table_anova %>% arrange(P_value)

```
We get results of association for all variables except utilities, which has two levels with one level only having one observation. The results of the various ANOVA and t-tests show that the categorical variables broadly show some significant association with LotFrontage, 34 categorical variables reach a traditionally accepted p < 0.05 significantly associated threshold. Despite this, most variables show an eta_squared value (percentage of total variance accounted for by differences between the groups) that would generally be described as small, the top 11 show an eta > 0.06, which is traditionally accepted as a moderate effect. In any case, a good model will be able to take into account these smaller effects when imputing these values.

ANOVA's have several assumptions that are key to the results being interpretable, no significant outliers, equality of variances, and normal distribution, and these should be tested before we accept these results. If groups are normally distributed with equal variances, then a normal ANOVA would be reasonable, however if variances are unequal, we need to apply Welch's correction for an ANOVA, and if neither assumption is met then we need to use the Games-Howell test to assess for association. 

The below function gathers the dependent and outcome variable, removing empty observations, skipping groups with 2 or fewer groups, combines groups with small sizes (i.e. < 5 observations) into a new group ("Other"), rechecks to see if there are a sufficient number of groups (> 2) and skips variables with insufficient groups. The function then checks normality and equality of variances, and then carries out the appropriate test of association for the group.

```{r}
library(dplyr)
library(car)
library(stats)

check_anova_assumptions_adjusted <- function(df, outcome_var, cat_vars,
                                             alpha = 0.05,
                                             min_per_group = 3,
                                             min_obs_per_level = 5,
                                             make_plots = FALSE) {
  results <- list()
  
  for (var in cat_vars) {
    cat("\n--- Checking:", var, "---\n")
    tmp <- df %>% 
      select(all_of(c(outcome_var, var))) %>%
      na.omit()
    
    #Only check groups we need to, i.e. those with more than 2 groups
    if (nlevels(tmp[[var]]) < 3) {
      cat("Skipped:", var, "has < 3 levels.\n")
      next
    }
    if (any(table(tmp[[var]]) < min_per_group)) {
      cat("Warning:", var, "has groups with fewer than", min_per_group, "samples.\n")
    }
    
    
    valid_levels <- names(which(table(df[[var]]) >= min_obs_per_level))
    tmp[[var]] <- forcats::fct_other(tmp[[var]],
                                 keep = valid_levels,
                                 other_level = "Other")
    tmp[[var]] <- droplevels(tmp[[var]])
    
    level_counts2 <- table(tmp[[var]])
    valid_levels2 <- names(level_counts2[level_counts2 >= min_obs_per_level])
    tmp <- tmp[tmp[[var]] %in% valid_levels2, , drop = FALSE]
    tmp[[var]] <- droplevels(tmp[[var]])
    
    if (nlevels(tmp[[var]]) < 3) {
      cat("Skipped:", var, "has < 3 levels after filtering for small groups.\n")
      next
    }
    
    
    
    formula_str <- paste(outcome_var, "~", var)
    aov_model <- tryCatch(aov(as.formula(formula_str), data = tmp),
                          error = function(e) NULL)
    if (is.null(aov_model)) next
    
    resid_vals <- residuals(aov_model)
    
    # Assumption tests
    shapiro_p <- tryCatch(shapiro.test(resid_vals)$p.value, error = function(e) NA)
    lev_p <- tryCatch(leveneTest(as.formula(formula_str), data = tmp)[1, "Pr(>F)"],
                      error = function(e) NA)
    
    # Choose appropriate test
    normality_ok <- !is.na(shapiro_p) && shapiro_p > alpha
    equal_var_ok <- !is.na(lev_p) && lev_p > alpha
    
    recommended <- if (normality_ok && equal_var_ok) {
      "Standard ANOVA"
    } else if (normality_ok && !equal_var_ok) {
      "Welch’s ANOVA"
    } else {
      "Games–Howell"
    }
    
    # Run selected test
    p_value <- NA
    if (recommended == "Standard ANOVA") {
      p_value <- tryCatch(summary(aov_model)[[1]][["Pr(>F)"]][1], error = function(e) NA)
    } else if (recommended == "Welch’s ANOVA") {
      welch <- tryCatch(oneway.test(as.formula(formula_str), data = tmp, var.equal = FALSE),
                        error = function(e) NULL)
      p_value <- if (!is.null(welch)) welch$p.value else NA
    } else if (recommended == "Games–Howell" && nlevels(tmp[[var]]) > 2) {
      gh <- tryCatch(games_howell_test(formula = as.formula(formula_str), data = tmp),
                     error = function(e) NULL)
      p_value <- if (!is.null(gh)) gh$p.adj else NA
    }
    
    cat("Normality:", ifelse(normality_ok, "OK", "Violated"),
        "| Equal variances:", ifelse(equal_var_ok, "OK", "Violated"),
        "| → Using:", recommended, "| p =", round(p_value, 5), "\n")
    
    if (make_plots) {
      par(mfrow = c(1, 2))
      qqnorm(resid_vals, main = paste("QQ Plot for", var))
      qqline(resid_vals)
      boxplot(as.formula(formula_str), data = tmp, main = paste("Boxplot for", var),
              col = "lightblue", las = 2)
    }
    
    results[[var]] <- data.frame(
      Variable = var,
      Levels = nlevels(tmp[[var]]),
      Shapiro_p = shapiro_p,
      Levene_p = lev_p,
      Test_p = p_value,
      Normality_OK = normality_ok,
      EqualVar_OK = equal_var_ok,
      Recommended_Test = recommended,
      stringsAsFactors = FALSE
    )
    }
  
  results_df <- do.call(rbind, results)
  rownames(results_df) <- NULL
  
  fisher_df <- results_df %>%
    group_by(Variable) %>%
    summarise(
      n_pairs = n(),
      P_value= fisher(Test_p)$p,
      df = 2*n_pairs,
    )
  
  return(fisher_df %>% arrange(P_value))
}

cat_vars <- names(train_csv)[sapply(train_csv, is.factor)]
results_table_gh <- check_anova_assumptions_adjusted(
  df = train_csv,
  outcome_var = "LotFrontage",
  cat_vars = cat_vars,
  alpha = 0.05,
  make_plots = FALSE
)

print(results_table_gh)
```

All variables that did test used the Games-Howell test, indicating that neither the variables are normally distributed nor were the variances between groups equal. Two variables started with two levels (Street and CentralAir) so were not going to be included in the ANOVA regardless, and two more simplified down to a two sample test (MiscFeature and PoolQC) once very rare groups were removed. This test provides a more stringent and more statistically rigorous method of testing for association between the dependent variable (in this case LotFrontage) and the categorical variables which shows 18 variables meet a traditional p < 0.05 threshold compared to the 34 found when using a standard ANOVA on all categorical variables, although for the least strict threshold I'll use (0.1) there are a total of 23 variables found to have some association. 

Whilst this does produce a list of variables that we are more confident about claiming an association between the dependent and outcome variable, does this list of variables actually perform better than the list produced by ANOVA analysis when imputing the unknown values? We'll perform some testing now to check if this is the case. The next two functions will be the same, with one using ANOVA testing solely to check which categorical variables are more closely related to the outcome and one using Games-Howell testing to check for association. Beyond this, both functions aim to do the same thing, generate a list of variables that are associated with the outcome (LotFrontage in this case), then test different prediction methods, linear regression, random forest, and chained equations (MICE), with a combination of different parameter values for these methods as part of cross validation.

Given it's a long function, and it can take a few minutes to run (or longer if the parameter grid is expanded), I'll explain briefly what it specifically does, as this will form the basis of the imputation of the other continuous variables, the sale price. Firstly, the parameter grid is set by the user, by making the alpha values more stringent (i.e. 1e-12) we can restrict the association to the top ~30 most associated variables, compared to 59 at the least stringent standards of p < 0.05. The function aims to evaluate each possible combination of parameters possible so be careful when expanding the grid as each addition does make runtime increase significantly. These combinations are ran in parallel using the plan and future packages for this reason. Variables are selected by association as previously described, data is masked using a few safety methods to ensure completeness and validity of testing, and then prediction is carried out using the previously noted methods.

```{r}
set.seed(123)

# Use your dataframe name
df_all <- train_csv

# PARAM GRID (customise as needed)
alpha_cor_values    <- c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
alpha_anova_values  <- c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
missforest_ntree    <- c(50, 100, 150, 200)
mice_methods        <- c("cart")        # keep small list to limit runtime

best_rmse <- NULL

# parallel plan: use most cores but leave one free
workers_to_use <- max(1, future::availableCores() - 1)
plan(multisession, workers = workers_to_use)

# make grid as a data.frame of combos
param_grid <- expand.grid(
  alpha_cor = alpha_cor_values,
  alpha_anova = alpha_anova_values,
  mf_ntree = missforest_ntree,
  mice_method = mice_methods,
  stringsAsFactors = FALSE
)

# The core evaluator
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 5,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 200) {

  # capture parameters
  alpha_cor   <- row_params$alpha_cor
  alpha_anova <- row_params$alpha_anova
  mf_ntree    <- row_params$mf_ntree
  mice_method <- row_params$mice_method
  
  # wrap in tryCatch so a single failing combo returns NA row
  tryCatch({
    df <- df  # use df_all by default
    
    # select numeric predictors by correlation 
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
   
    # select factor predictors by ANOVA 
    selected_factors <- results_table_anova$Variable[results_table_anova$P_value < alpha_anova]
    
    selected_predictors <- c(selected_numeric, selected_factors)
    if (length(selected_predictors) == 0) {
      # Return parameter settings
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_anova))
    }
    
    # keep only rows where predictors are complete (fair comparison) --
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]
    
    # known indices (where target known)
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      # not enough data to evaluate this combo properly -> return NA results
      return(tibble(
        alpha_cor = alpha_cor, alpha_anova = alpha_anova,
        en_alpha = en_alpha, mf_ntree = mf_ntree, mice_method = mice_method,
        Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = length(selected_predictors)
      ))
    }
    
    # mask creation (ensure training retains factor levels) --
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL
    
    # helper for mask validity
    # checks there are enough levels in the factors present 
    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]
    
    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      # fallback try smaller mask fractions
      found <- FALSE
      for (mf in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf))
        attempt2 <- 1
        while(attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
    }
    
    # Prepare masked dataframe 
    df_model_masked <- model_df
    true_values <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA  # mask
    
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final
    
    # Drop single-level factors or constant numerics in training
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) {
            final_predictors <- setdiff(final_predictors, f)
          }
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) {
          final_predictors <- setdiff(final_predictors, nvar)
        }
      }
    }
    
    # Build train/test frames
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]
    
    # Align factor levels: set test levels to training levels (prevents "new levels" error)
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)  # unseen => NA (safe)
      }
    }
    
    # -------- Linear regression --------
    lm_pred <- rep(NA_real_, length(test_idx))
    lm_success <- FALSE
    if (nrow(train_df_reg) >= 5 && length(final_predictors) > 0) {
      lm_formula <- as.formula(paste(target_var, "~", paste(final_predictors, collapse = " + ")))
      lm_fit <- tryCatch(lm(lm_formula, data = train_df_reg), error = function(e) e)
      if (!inherits(lm_fit, "error")) {
        # predict; rows with NA predictors will produce NA predictions
        lm_pred <- tryCatch(predict(lm_fit, newdata = test_df_reg), error = function(e) rep(NA_real_, nrow(test_df_reg)))
        lm_success <- TRUE
      }
    }
    
    # -------- MICE (masking already done in test_df_reg: LotFrontage is NA in test rows) --------
    mice_pred <- rep(NA_real_, length(test_idx))
    mice_success <- FALSE
    # Combine train+test (with test's LotFrontage currently NA) and impute
    combined_for_mice <- rbind(train_df_reg, test_df_reg)
    mice_result <- tryCatch(mice(combined_for_mice, m = 1, maxit = 5, method = mice_method, printFlag = FALSE),
                            error = function(e) NULL)
    if (!is.null(mice_result)) {
      mice_imputed <- complete(mice_result)
      ntr <- nrow(train_df_reg)
      mice_test  <- mice_imputed[(ntr + 1):nrow(mice_imputed), , drop = FALSE]
      mice_pred <- mice_test[[target_var]]
      mice_success <- TRUE
    }
    
    # -------- missForest (masking already done in test_df_reg) --------
    mf_pred <- rep(NA_real_, length(test_idx))
    mf_success <- FALSE
    combined_for_mf <- rbind(train_df_reg, test_df_reg)
    # ensure types are appropriate; missForest expects factors as factors and numerics numeric
    mf_out <- tryCatch(missForest(combined_for_mf, maxiter = 10, ntree = mf_ntree, verbose = TRUE),
                       error = function(e) NULL)
    if (!is.null(mf_out)) {
      mf_imputed <- mf_out$ximp
      ntr <- nrow(train_df_reg)
      mf_pred <- mf_imputed[[target_var]][(ntr + 1):nrow(mf_imputed)]
      mf_success <- TRUE
    }
    
    
    # Evaluate metrics using the true_values vector and each predicted vector
    eval_metrics <- function(true, pred) {
      ok <- !is.na(true) & !is.na(pred)
      if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
      t <- true[ok]; p <- pred[ok]
      list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
    }
    
    
    res_lm   <- eval_metrics(true_values, lm_pred)
    res_mice <- eval_metrics(true_values, mice_pred)
    res_mf   <- eval_metrics(true_values, mf_pred)

    # Compose results tibble for this combination
    result_tib <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "MICE", "missForest"),
      RMSE = c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE),
      MAE  = c(res_lm$MAE, res_mice$MAE, res_mf$MAE),
      R2   = c(res_lm$R2, res_mice$R2, res_mf$R2),
      n_predictors = length(final_predictors)
    )
    
    return(result_tib)
    
  }, error = function(e) {
    # On error, return NA result with the parameters
    tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_,
      n_predictors = NA_integer_
    )
  })
}

# Run grid in parallel (use reproducible random seeds)
# convert each row of param_grid to a list for mapping
param_list <- split(param_grid, seq_len(nrow(param_grid)))

# Use future_map_dfr with reproducible seeds
results_all_anova <- future_map_dfr(param_list, evaluate_combo, .options = furrr_options(seed = TRUE))

# Clean up: restore sequential plan if desired
plan(sequential)

# You can also inspect best combos:
best_by_method <- results_all_anova %>%
  group_by(Method) %>%
  filter(!is.na(RMSE)) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup()
print(best_by_method)
```
```{r}
set.seed(123)

# Use your dataframe name
df_all <- train_csv

# PARAM GRID (customise as needed)
alpha_cor_values    <- c(1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
alpha_gh_values     <- c(1e-12, 1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2, 1e-1)
missforest_ntree    <- c(10, 25, 50, 100, 150, 200, 250, 300, 350, 400)
mice_methods        <- c("cart")        # keep small list to limit runtime

# parallel plan: use most cores but leave one free
workers_to_use <- max(1, future::availableCores() - 1)
plan(multisession, workers = workers_to_use)

# make grid as a data.frame of combos
param_grid <- expand.grid(
  alpha_cor = alpha_cor_values,
  alpha_gh = alpha_gh_values,
  mf_ntree = missforest_ntree,
  mice_method = mice_methods,
  stringsAsFactors = FALSE
)

# The core evaluator
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 10,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 20) {
  # capture parameters
  alpha_cor   <- row_params$alpha_cor
  alpha_gh    <- row_params$alpha_gh
  mf_ntree    <- row_params$mf_ntree
  mice_method <- row_params$mice_method
  
  # wrap in tryCatch so a single failing combo returns NA row
  tryCatch({
    df <- df  # use df_all by default
    
    # select numeric predictors by correlation 
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
   
    # select factor predictors by Games-Howell 
    selected_factors <- results_table_gh$Variable[results_table_gh$P_value < alpha_gh]
    
    selected_predictors <- c(selected_numeric, selected_factors)
    if (length(selected_predictors) == 0) {
      # Return parameter settings
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_gh))
    }
    
    
      # keep only rows where predictors are complete 
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]
    
    # known indices (where target known)
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      # not enough data to evaluate this combo properly -> return NA results
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_gh))
    }
    
    # mask creation (ensure training retains factor levels) --
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL
    
    # helper for mask validity
    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]
    
    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      # fallback try smaller mask fractions
      found <- FALSE
      for (mf in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf))
        attempt2 <- 1
        while(attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
      if (!found) {
        return(sprintf('No masking found at masking fraction %d and max mask attempts %f, please adjust these values and try again', mask_frac, max_mask_attempts))
      }
    }
    
    # Prepare masked dataframe (same approach as your working code)
    df_model_masked <- model_df
    true_values <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA  # mask
    
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final
    
    # Drop single-level factors or constant numerics in training
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) {
            final_predictors <- setdiff(final_predictors, f)
          }
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) {
          final_predictors <- setdiff(final_predictors, nvar)
        }
      }
    }
    if (length(final_predictors) == 0) {
      return(tibble(
        alpha_cor = alpha_cor, alpha_gh = alpha_gh,
        en_alpha = en_alpha, mf_ntree = mf_ntree, mice_method = mice_method,
        Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = 0
      ))
    }
    
    # Build train/test frames
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]
    
    # Align factor levels: set test levels to training levels (prevents "new levels" error)
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)  # unseen => NA (safe)
      }
    }
    
    # -------- Linear regression --------
    lm_pred <- rep(NA_real_, length(test_idx))
    lm_success <- FALSE
    if (nrow(train_df_reg) >= 5 && length(final_predictors) > 0) {
      lm_formula <- as.formula(paste(target_var, "~", paste(final_predictors, collapse = " + ")))
      lm_fit <- tryCatch(lm(lm_formula, data = train_df_reg), error = function(e) e)
      if (!inherits(lm_fit, "error")) {
        # predict; rows with NA predictors will produce NA predictions
        lm_pred <- tryCatch(predict(lm_fit, newdata = test_df_reg), error = function(e) rep(NA_real_, nrow(test_df_reg)))
        lm_success <- TRUE
      }
    }
    
    # -------- MICE (masking already done in test_df_reg: LotFrontage is NA in test rows) --------
    mice_pred <- rep(NA_real_, length(test_idx))
    mice_success <- FALSE
    # Combine train+test (with test's LotFrontage currently NA) and impute
    combined_for_mice <- rbind(train_df_reg, test_df_reg)
    mice_result <- tryCatch(mice(combined_for_mice, m = 1, maxit = 5, method = mice_method, printFlag = FALSE),
                            error = function(e) NULL)
    if (!is.null(mice_result)) {
      mice_imputed <- complete(mice_result)
      ntr <- nrow(train_df_reg)
      mice_test  <- mice_imputed[(ntr + 1):nrow(mice_imputed), , drop = FALSE]
      mice_pred <- mice_test[[target_var]]
      mice_success <- TRUE
    }
    
    # -------- missForest (masking already done in test_df_reg) --------
    mf_pred <- rep(NA_real_, length(test_idx))
    mf_success <- FALSE
    combined_for_mf <- rbind(train_df_reg, test_df_reg)
    # ensure types are appropriate; missForest expects factors as factors and numerics numeric
    mf_out <- tryCatch(missForest(combined_for_mf, maxiter = 10, ntree = mf_ntree, verbose = FALSE),
                       error = function(e) NULL)
    if (!is.null(mf_out)) {
      mf_imputed <- mf_out$ximp
      ntr <- nrow(train_df_reg)
      mf_pred <- mf_imputed[[target_var]][(ntr + 1):nrow(mf_imputed)]
      mf_success <- TRUE
    }
    
    # Evaluate metrics using the true_values vector and each predicted vector
    eval_metrics <- function(true, pred) {
      ok <- !is.na(true) & !is.na(pred)
      if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
      t <- true[ok]; p <- pred[ok]
      list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
    }
    
    res_lm   <- eval_metrics(true_values, lm_pred)
    res_mice <- eval_metrics(true_values, mice_pred)
    res_mf   <- eval_metrics(true_values, mf_pred)
    
    # Compose results tibble for this combination
    result_tib <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_gh = alpha_gh,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "MICE", "missForest"),
      RMSE = c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE),
      MAE  = c(res_lm$MAE, res_mice$MAE, res_mf$MAE),
      R2   = c(res_lm$R2, res_mice$R2, res_mf$R2),
      n_predictors = length(final_predictors)
    )
    
    return(result_tib)
    
  }, error = function(e) {
    # On error, return NA result with the parameters
    tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_gh = alpha_gh,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_,
      n_predictors = NA_integer_
    )
  })
}

# Run grid in parallel (use reproducible random seeds)
# convert each row of param_grid to a list for mapping
param_list <- split(param_grid, seq_len(nrow(param_grid)))

# Use future_map_dfr with reproducible seeds
results_all_gh <- future_map_dfr(param_list, evaluate_combo, .options = furrr_options(seed = TRUE))

# Clean up: restore sequential plan if desired
plan(sequential)

# You can also inspect best combos:
best_by_method <- results_all_gh %>%
  group_by(Method) %>%
  filter(!is.na(RMSE)) %>%
  slice_min(RMSE, n = 1, with_ties = FALSE) %>%
  ungroup()
print(best_by_method)
```
Again we see that the best predictions are found by combining as many predictors as we can (in this case 25 predictors) and using a random forest prediction model, achieving a RMSE of 11.42, MAE of 7.64, and R2 of 0.69. This is not much worse than the results found by using ANOVA alone to generate categorical variables associated with the outcome variable, but it is worse than the results found using ANOVA alone to associate categorical variables (RMSE = 10.28, R2 = 0.72). 

Deep learning models generally don't perform better than random forests on tabular imputation datasets like this, but just to check I've included one here. More for my sanity rather than any statistical completeness.

```{r}
# Load packages
library(dplyr)
library(forcats)
library(tidyr)
library(h2o)
library(rstatix)
library(tibble)

# Initialize H2O once
h2o.init(nthreads = 1, max_mem_size = "8G")

# ---- CONFIG ----
set.seed(123)
df_all <- train_csv
target_var <- "LotFrontage"

alpha_cor <- 0.01
alpha_gh  <- 0.01
mask_frac <- 0.2

# ---- Function ----
evaluate_h2o_deep <- function(df,
                              target_var = "LotFrontage",
                              alpha_cor = 0.01,
                              alpha_gh = 0.01,
                              mask_frac = 0.2,
                              min_obs_cor = 10,
                              min_obs_per_level = 5,
                              min_obs_anova = 20) {
  
  # --- Step 0: Clean and convert types ---
  df <- df %>%
    mutate(across(where(is.ordered), ~ factor(.x, ordered = FALSE))) %>%  # fix ordered factors
    mutate(across(where(is.character), as.factor))                        # ensure characters are factors
  
  # Identify variable types
  numeric_vars <- names(df)[sapply(df, is.numeric) & names(df) != target_var]
  factor_vars  <- names(df)[sapply(df, is.factor)]

  # --- Step 1: Select numeric predictors by correlation ---
  selected_numeric <- c()
  for (v in numeric_vars) {
    idx <- which(!is.na(df[[target_var]]) & !is.na(df[[v]]))
    if (length(idx) < min_obs_cor) next
    ct <- tryCatch(cor.test(df[[target_var]][idx], df[[v]][idx]), error = function(e) NULL)
    if (!is.null(ct) && !is.na(ct$p.value) && ct$p.value < alpha_cor) {
      selected_numeric <- c(selected_numeric, v)
    }
  }

  # --- Step 2: Select categorical predictors by Games-Howell test ---
  selected_factors <- c()
  for (v in factor_vars) {
    tmp <- df %>% select(all_of(c(target_var, v))) %>% drop_na()
    if (nrow(tmp) < min_obs_anova) next

    # Collapse rare levels
    valid_levels <- names(which(table(tmp[[v]]) >= min_obs_per_level))
    tmp[[v]] <- fct_other(tmp[[v]], keep = valid_levels, other_level = "Other")
    tmp[[v]] <- droplevels(tmp[[v]])

    if (nlevels(tmp[[v]]) < 2) next

    gh <- tryCatch(games_howell_test(as.formula(paste(target_var, "~", v)), data = tmp),
                   error = function(e) NULL)
    if (is.null(gh)) next

    p_values <- gh$p.adj
    p_values <- p_values[!is.na(p_values)]
    if (length(p_values) == 0) next
    chi_sq <- -2 * sum(log(p_values))
    combined_p <- 1 - pchisq(chi_sq, df = 2 * length(p_values))
    if (combined_p < alpha_gh) selected_factors <- c(selected_factors, v)
  }

  predictors <- c(selected_numeric, selected_factors)
  if (length(predictors) == 0) stop("No predictors selected; adjust alpha thresholds.")

  # --- Step 3: Prepare modeling data ---
  model_df <- df %>% select(all_of(c(target_var, predictors))) %>% drop_na()

  # Split known values
  known_idx <- which(!is.na(model_df[[target_var]]))
  mask_n <- max(1, round(length(known_idx) * mask_frac))
  mask_idx <- sample(known_idx, mask_n)

  train_df <- model_df[-mask_idx, ]
  test_df  <- model_df[mask_idx, ]

  true_values <- test_df[[target_var]]

  # --- Step 4: Convert to H2O frames ---
  train_h2o <- as.h2o(train_df)
  test_h2o  <- as.h2o(test_df)

  predictors <- setdiff(names(train_df), target_var)

  # --- Step 5: Deep Learning model ---
  grid_id <- paste0("dl_grid_", sample(1000:9999, 1))
  hyper_params <- list(
    hidden = list(c(32, 16), c(64, 32, 16), c(128, 64)),
    activation = c("Rectifier", "Tanh", "Maxout"),
    input_dropout_ratio = c(0, 0.1, 0.2),
    rate = c(0.01, 0.005, 0.001)
  )

  dl_grid <- h2o.grid(
    algorithm = "deeplearning",
    grid_id = grid_id,
    x = predictors,
    y = target_var,
    training_frame = train_h2o,
    validation_frame = test_h2o,
    hyper_params = hyper_params,
    epochs = 100,
    stopping_metric = "RMSE",
    stopping_rounds = 5,
    stopping_tolerance = 1e-4,
    seed = 123,
    standardize = TRUE
  )

  # --- Step 6: Best model and performance ---
  grid_perf <- h2o.getGrid(grid_id, sort_by = "RMSE", decreasing = FALSE)
  best_model_id <- grid_perf@model_ids[[1]]
  best_dl <- h2o.getModel(best_model_id)

  preds <- as.data.frame(h2o.predict(best_dl, test_h2o))[, 1]

  ok <- !is.na(true_values) & !is.na(preds)
  RMSE <- sqrt(mean((true_values[ok] - preds[ok])^2))
  MAE  <- mean(abs(true_values[ok] - preds[ok]))
  R2   <- cor(true_values[ok], preds[ok])^2

  tibble(
    Method = "H2O Deep Learning",
    alpha_cor = alpha_cor,
    alpha_gh = alpha_gh,
    RMSE = RMSE,
    MAE = MAE,
    R2 = R2,
    n_predictors = length(predictors)
  )
}

# ---- RUN ----
results_h2o <- evaluate_h2o_deep(df_all, alpha_cor = 0.01, alpha_gh = 0.01)
print(results_h2o)

# ---- CLEANUP ----
h2o.shutdown(prompt = FALSE)
```

Good for confirming my sanity, the deep learning model is performing worse on predicting this data when compared to a random forest model. 

Now we have our optimal prediction model and optimal parameters for this model, lets try to optimise it further. We'll look at the results that are particularly inaccurate and try to see if there are any patterns to those.


```{r}
set.seed(123)

# Use your dataframe name
df_all <- train_csv

# PARAM GRID (customise as needed)
alpha_cor_values    <- c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
alpha_anova_values  <- c(1e-10, 1e-8, 1e-6, 1e-4, 1e-2, 5e-2)
missforest_ntree    <- c(50, 100, 150, 200)
mice_methods        <- c("cart")        # keep small list to limit runtime

collapse_rare_levels <- function(df, factor_vars, min_obs ) {
  for (v in factor_vars) {
    if (!is.factor(df[[v]])) next

    tbl <- table(df[[v]])
    rare_levels <- names(tbl)[tbl < min_obs]

    if (length(rare_levels) > 0) {
      df[[v]] <- forcats::fct_collapse(
        df[[v]],
        Other = rare_levels
      )
      df[[v]] <- droplevels(df[[v]])
    }
  }
  df
}

# parallel plan: use most cores but leave one free
workers_to_use <- max(1, future::availableCores() - 1)
plan(multisession, workers = workers_to_use)

# make grid as a data.frame of combos
param_grid <- expand.grid(
  alpha_cor = alpha_cor_values,
  alpha_anova = alpha_anova_values,
  mf_ntree = missforest_ntree,
  mice_method = mice_methods,
  stringsAsFactors = FALSE
)

# The core evaluator
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 10,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 200) {

  # capture parameters
  alpha_cor   <- row_params$alpha_cor
  alpha_anova <- row_params$alpha_anova
  mf_ntree    <- row_params$mf_ntree
  mice_method <- row_params$mice_method
  
  # wrap in tryCatch so a single failing combo returns NA row
  tryCatch({
    df <- df  # use df_all by default
    
    # Identify factor variables among selected predictors
    all_factor_vars <- names(df)[sapply(df, is.factor)]

    # Collapse rare levels BEFORE selecting predictor-complete rows
    df <- collapse_rare_levels(df, all_factor_vars, min_obs = min_obs_per_level)
    
    # select numeric predictors by correlation 
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
   
    # select factor predictors by ANOVA 
    selected_factors <- results_table_anova$Variable[results_table_anova$P_value < alpha_anova]
    
    selected_predictors <- c(selected_numeric, selected_factors)
    if (length(selected_predictors) == 0) {
      # Return parameter settings
      return(sprintf('No predictors found at p_values of %d for correlation and %s for ANOVA. Please adjust these values and try again', alpha_cor, alpha_anova))
    }
    
    # keep only rows where predictors are complete (fair comparison) --
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]
    
    
    
    # known indices (where target known)
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      # not enough data to evaluate this combo properly -> return NA results
      return(tibble(
        alpha_cor = alpha_cor, alpha_anova = alpha_anova,
        en_alpha = en_alpha, mf_ntree = mf_ntree, mice_method = mice_method,
        Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = length(selected_predictors)
      ))
    }
    
    # mask creation (ensure training retains factor levels) --
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL
    
    # helper for mask validity
    # checks there are enough levels in the factors present 
    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]
    
    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      # fallback try smaller mask fractions
      found <- FALSE
      for (mf in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf))
        attempt2 <- 1
        while(attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
    }
    
    # Prepare masked dataframe 
    df_model_masked <- model_df
    true_values <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA  # mask
    
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final
    
    # Drop single-level factors or constant numerics in training
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) {
            final_predictors <- setdiff(final_predictors, f)
          }
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) {
          final_predictors <- setdiff(final_predictors, nvar)
        }
      }
    }
    
    # Build train/test frames
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]
    
    # Align factor levels: set test levels to training levels (prevents "new levels" error)
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)  # unseen => NA (safe)
      }
    }
    
    # -------- Linear regression --------
    lm_pred <- rep(NA_real_, length(test_idx))
    lm_success <- FALSE
    if (nrow(train_df_reg) >= 5 && length(final_predictors) > 0) {
      lm_formula <- as.formula(paste(target_var, "~", paste(final_predictors, collapse = " + ")))
      lm_fit <- tryCatch(lm(lm_formula, data = train_df_reg), error = function(e) e)
      if (!inherits(lm_fit, "error")) {
        # predict; rows with NA predictors will produce NA predictions
        lm_pred <- tryCatch(predict(lm_fit, newdata = test_df_reg), error = function(e) rep(NA_real_, nrow(test_df_reg)))
        lm_success <- TRUE
      }
    }
    
    # -------- MICE (masking already done in test_df_reg: LotFrontage is NA in test rows) --------
    mice_pred <- rep(NA_real_, length(test_idx))
    mice_success <- FALSE
    # Combine train+test (with test's LotFrontage currently NA) and impute
    combined_for_mice <- rbind(train_df_reg, test_df_reg)
    mice_result <- tryCatch(mice(combined_for_mice, m = 1, maxit = 5, method = mice_method, printFlag = FALSE),
                            error = function(e) NULL)
    if (!is.null(mice_result)) {
      mice_imputed <- complete(mice_result)
      ntr <- nrow(train_df_reg)
      mice_test  <- mice_imputed[(ntr + 1):nrow(mice_imputed), , drop = FALSE]
      mice_pred <- mice_test[[target_var]]
      mice_success <- TRUE
    }
    
    # -------- missForest (masking already done in test_df_reg) --------
    mf_pred <- rep(NA_real_, length(test_idx))
    mf_success <- FALSE
    combined_for_mf <- rbind(train_df_reg, test_df_reg)
    # ensure types are appropriate; missForest expects factors as factors and numerics numeric
    mf_out <- tryCatch(missForest(combined_for_mf, maxiter = 10, ntree = mf_ntree, verbose = TRUE),
                       error = function(e) NULL)
    if (!is.null(mf_out)) {
      mf_imputed <- mf_out$ximp
      ntr <- nrow(train_df_reg)
      mf_pred <- mf_imputed[[target_var]][(ntr + 1):nrow(mf_imputed)]
      mf_success <- TRUE
    }
    
    # --- STORE ALL RAW PREDICTIONS FOR LATER ANALYSIS ---
  all_predictions <- tibble::tibble(
    index = mask_indices_final,
    true = true_values,
    pred_lm = lm_pred,
    pred_mice = mice_pred,
    pred_mf = mf_pred
  )
    
    # Evaluate metrics using the true_values vector and each predicted vector
    eval_metrics <- function(true, pred) {
      ok <- !is.na(true) & !is.na(pred)
      if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
      t <- true[ok]; p <- pred[ok]
      list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
    }
    
    
    res_lm   <- eval_metrics(true_values, lm_pred)
    res_mice <- eval_metrics(true_values, mice_pred)
    res_mf   <- eval_metrics(true_values, mf_pred)
    # Determine best model by RMSE
    rmse_vec <- c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE)
    names(rmse_vec) <- c("lm", "mice", "mf")

    best_method <- names(which.min(rmse_vec))
    best_pred_vector <- switch(best_method,
                           lm = lm_pred,
                           mice = mice_pred,
                           mf = mf_pred)

    best_prediction <- tibble::tibble(
      index = mask_indices_final,
      true = true_values,
      pred = best_pred_vector,
      abs_error = abs(true_values - best_pred_vector),
      method = best_method
  )

    # Compose results tibble for this combination
    result_tib <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "MICE", "missForest"),
      RMSE = c(res_lm$RMSE, res_mice$RMSE, res_mf$RMSE),
      MAE  = c(res_lm$MAE, res_mice$MAE, res_mf$MAE),
      R2   = c(res_lm$R2, res_mice$R2, res_mf$R2),
      n_predictors = length(final_predictors)
    )
    
    return(list(
      test_df = test_df_reg,
      metrics_table = result_tib,
      all_predictions = all_predictions,
      best_prediction = best_prediction,
      n_predictors = length(final_predictors)
    ))
    
  }, error = function(e) {
    # On error, return NA result with the parameters
    tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      mice_method = mice_method,
      Method = c("Linear Regression", "Elastic Net", "MICE", "missForest"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_,
      n_predictors = NA_integer_
    )
  })
}

# Run grid in parallel (use reproducible random seeds)
# convert each row of param_grid to a list for mapping
param_list <- split(param_grid, seq_len(nrow(param_grid)))

# 1) Run the grid producing a list of results (each element is a list returned by evaluate_combo)
results_list <- future_map(param_list, evaluate_combo, .options = furrr_options(seed = TRUE))

# 2) Remove any NULL or error-return entries safely
# (If evaluate_combo returned a tibble on error, adjust the filter accordingly)
valid_results <- keep(results_list, ~ !is.null(.x) && is.list(.x))

# 3) Extract metrics: this binds all metrics_table tibbles into one long tibble
metrics_all <- map_dfr(valid_results, function(res) {
  # res$metrics_table should be a tibble (possibly with multiple rows: one per method)
  if (is.null(res$metrics_table)) return(NULL)
  # ensure it includes the parameter columns (alpha_cor etc.)
  as_tibble(res$metrics_table)
}, .id = "combo_id")   # optional combo id if you want



# Now you have:
# metrics_all        : tibble of metrics (many rows; each row = one method entry returned by evaluate_combo)


# Example: inspect top-performing combos by RMSE
best_by_combo <- metrics_all %>%
  group_by(alpha_cor, alpha_anova, mf_ntree, mice_method, Method) %>%
  summarise(RMSE = mean(RMSE, na.rm = TRUE), .groups = "drop") %>%
  arrange(RMSE) %>%
  slice_head(n = 20)

print(best_by_combo)



```

Collapsing rare levels does provide an improvement in RMSE, but only a very small one (10.28 vs 10.23). Still, we now have a list of best predictions, so we'll examine the predictions there that were particularly inaccurate and see if any trend is possible for identifying the subset of inaccurate predictions. Even looking at those results which were more than one SD greater than the mean error, we're only examining the worst 10% of predictions.

```{r}
best_prediction <- valid_results[[56]]$best_prediction


histogram(best_prediction$abs_error)
analyse_bad_predictions <- function(df_all,
                                    predictions,
                                    true_values,
                                    predictors,
                                    n_sd = 1,
                                    p_thresh = 0.05) {
  suppressPackageStartupMessages({
    library(dplyr)
    library(tibble)
    library(stats)
  })

  # ---------------------------
  # 1. Compute prediction error
  # ---------------------------
  err <- predictions - true_values
  mu  <- mean(err, na.rm = TRUE)
  sdv <- sd(err, na.rm = TRUE)

  bad_idx <- which(err > mu + n_sd * sdv)

  if (length(bad_idx) == 0) {
    message("No bad predictions detected.")
    return(NULL)
  }

  bad_data  <- df_all[bad_idx, predictors, drop = FALSE]
  good_data <- df_all[-bad_idx, predictors, drop = FALSE]

  # ---------------------------
  # 2. Numeric predictors
  # ---------------------------
  numeric_vars <- names(bad_data)[sapply(bad_data, is.numeric)]
  numeric_summary <- NULL

  if (length(numeric_vars) > 0) {
    numeric_summary <- lapply(numeric_vars, function(v) {
      bad_mean  <- mean(bad_data[[v]], na.rm = TRUE)
      good_mean <- mean(good_data[[v]], na.rm = TRUE)
      tibble(
        Variable = v,
        Mean_Bad = bad_mean,
        Mean_Good = good_mean,
        Mean_Diff = bad_mean - good_mean
      )
    }) |> bind_rows()
  }

  # ---------------------------
  # 3. Factor predictors
  # ---------------------------
  factor_vars <- names(bad_data)[sapply(bad_data, is.factor)]
  factor_summary <- list()

  for (v in factor_vars) {
    full_tab <- table(df_all[[v]])
    bad_tab  <- table(bad_data[[v]])
    good_tab <- table(good_data[[v]])

    full_prop <- full_tab / sum(full_tab)
    bad_prop  <- bad_tab / sum(bad_tab)

    all_levels <- union(names(full_tab), names(bad_tab))
    full_prop <- full_prop[all_levels]
    bad_prop  <- bad_prop[all_levels]
    full_prop[is.na(full_prop)] <- 0
    bad_prop[is.na(bad_prop)] <- 0

    over_rep <- bad_prop - full_prop

    # ---- Statistical significance ----
    p_values <- sapply(all_levels, function(lvl) {
      # Construct 2x2 table for chi-squared or Fisher test
      bad_yes <- sum(bad_data[[v]] == lvl, na.rm = TRUE)
      bad_no  <- nrow(bad_data) - bad_yes
      good_yes <- sum(good_data[[v]] == lvl, na.rm = TRUE)
      good_no  <- nrow(good_data) - good_yes

      tbl <- matrix(c(bad_yes, bad_no, good_yes, good_no), nrow = 2, byrow = TRUE)

      # Use Fisher's test if any expected count < 5
      if (any(tbl < 5)) {
        test <- tryCatch(fisher.test(tbl)$p.value, error = function(e) NA)
      } else {
        test <- tryCatch(chisq.test(tbl, correct = FALSE)$p.value, error = function(e) NA)
      }
      test
    })

    factor_summary[[v]] <- tibble(
      Variable = v,
      Level = all_levels,
      Prop_Bad = as.numeric(bad_prop),
      Prop_All = as.numeric(full_prop),
      Over_Rep = as.numeric(over_rep),
      P_value = p_values
    )
  }

  # ---------------------------
  # 4. Generate text summary
  # ---------------------------
  text_lines <- c()

  # Numeric summary
  if (!is.null(numeric_summary)) {
    big_shifts <- numeric_summary %>%
      arrange(desc(abs(Mean_Diff))) %>%
      head(3)

    for (i in seq_len(nrow(big_shifts))) {
      txt <- sprintf(
        "- %s differs between bad and good predictions (Δ = %.2f)",
        big_shifts$Variable[i], big_shifts$Mean_Diff[i]
      )
      text_lines <- c(text_lines, txt)
    }
  }

  # Factor summary (only significant over-representations)
  for (v in names(factor_summary)) {
    tab <- factor_summary[[v]] %>%
      filter(P_value < p_thresh) %>%
      arrange(desc(Over_Rep))

    if (nrow(tab) > 0) {
      top <- head(tab, 1)
      txt <- sprintf(
        "- Factor %s: Level '%s' is significantly over-represented among bad predictions (%.2f%% vs %.2f%%, p = %.3f)",
        top$Variable[1],
        top$Level[1],
        100 * top$Prop_Bad[1],
        100 * top$Prop_All[1],
        top$P_value[1]
      )
      text_lines <- c(text_lines, txt)
    }
  }

  # ---------------------------
  # Final return
  # ---------------------------
  list(
    bad_indices = bad_idx,
    bad_data = bad_data,
    numeric_summary = numeric_summary,
    factor_summary = factor_summary,
    text_summary = paste(text_lines, collapse = "\n")
  )
}

   
diagnostic <- analyse_bad_predictions(df_all, predictions = best_prediction$pred, true_values = best_prediction$true, predictors = names(df_all), n_sd = 1, p_thresh = 0.1)
cat(diagnostic$text_summary)
```
We can see a few variables which are over-represented in the inaccurate predictions. SalePrice and ID are clearly not useful so will not be used, but other variables (LotArea, LandContour, Neighborhood, RoofStyle, Exterior1st, Exterior2nd, BsmtExposure, BsmtFinType1, Electrical, Fence, MoSold) are useful variables to be used in diagnosing the bad predictions. Let's further examine these inaccurate predictions.
```{r}
n_sd <- 0.5
mean_error <- mean(best_prediction$abs_error)
sd_error <- sd(best_prediction$abs_error)
threshold <- mean_error + (n_sd * sd_error)
best_prediction$is_bad <- best_prediction$abs_error > threshold

good_prediction <- subset(best_prediction, is_bad == FALSE)
bad_prediction <- subset(best_prediction, is_bad == TRUE)
```
Here we can see that there is a very small number of very bad predictions, with n_sd = 3, there are 8 bad predictions, with n_sd = 2, there are 13 bad predictions, with n_sd = 1, there are 29 bad predictions, with n_sd = 0.5, there are 43 bad predictions. We'll try to look at these results specifically and see how we can improve these predictions. 

We'll first try different models for these predictions
```{r}
evaluate_combo <- function(row_params, df = df_all,
                           target_var = "LotFrontage",
                           mask_frac = 0.20,
                           min_obs_per_level = 10,
                           min_obs_cor = 10,
                           min_obs_anova = 20,
                           min_rows_train = 30,
                           max_mask_attempts = 200,
                           cv_folds = 5,
                           hard_resid_sd_mult = 1.5,
                           specialist_min_rows = 20,
                           specialist_prob_threshold = 0.5,
                           mf_ntree = 100,
                           xgb_nrounds = 100,
                           verbose = FALSE) {
  # Required packages
  requireNamespace("dplyr"); requireNamespace("forcats"); requireNamespace("tibble")
  requireNamespace("purrr"); requireNamespace("missForest"); requireNamespace("xgboost")
  requireNamespace("randomForest")
  library(dplyr); library(forcats); library(tibble); library(purrr)

  # helpers -------------------------------------------------------------------

  # safe rbind that aligns columns and factor levels
  safe_rbind <- function(a, b) {
    all_cols <- union(names(a), names(b))
    add_missing <- function(df, cols) {
      for (cc in cols) if (!(cc %in% names(df))) df[[cc]] <- NA
      df[, cols, drop = FALSE]
    }
    a2 <- add_missing(a, all_cols)
    b2 <- add_missing(b, all_cols)
    # align factor levels
    for (col in all_cols) {
      if (is.factor(a2[[col]]) || is.factor(b2[[col]])) {
        levs <- union(if (is.factor(a2[[col]])) levels(a2[[col]]) else NULL,
                      if (is.factor(b2[[col]])) levels(b2[[col]]) else NULL)
        if (is.factor(a2[[col]])) a2[[col]] <- factor(a2[[col]], levels = levs) else a2[[col]] <- factor(a2[[col]], levels = levs)
        if (is.factor(b2[[col]])) b2[[col]] <- factor(b2[[col]], levels = levs) else b2[[col]] <- factor(b2[[col]], levels = levs)
      }
    }
    rbind(a2, b2)
  }

  # collapse rare levels in specified factor vars (used AFTER masking)
  collapse_rare_levels <- function(df_local, factor_vars, min_obs) {
    for (v in factor_vars) {
      if (! (v %in% names(df_local))) next
      if (!is.factor(df_local[[v]])) next
      tbl <- table(df_local[[v]])
      rare_levels <- names(tbl)[tbl < min_obs]
      if (length(rare_levels) > 0) {
        df_local[[v]] <- forcats::fct_collapse(df_local[[v]], Other = rare_levels)
        df_local[[v]] <- droplevels(df_local[[v]])
      }
    }
    df_local
  }

  # rmse/mae/r2 evaluator
  eval_metrics <- function(true, pred) {
    ok <- !is.na(true) & !is.na(pred)
    if (sum(ok) == 0) return(list(n = 0, RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_))
    t <- true[ok]; p <- pred[ok]
    list(n = sum(ok), RMSE = sqrt(mean((t - p)^2)), MAE = mean(abs(t - p)), R2 = if (length(t) > 1) cor(t, p)^2 else NA_real_)
  }

  # parameters from row
  alpha_cor   <- row_params$alpha_cor %||% row_params$alpha_cor
  alpha_anova <- row_params$alpha_anova %||% row_params$alpha_anova
  # allow mf_ntree from row_params too
  if (!is.null(row_params$mf_ntree)) mf_ntree <- row_params$mf_ntree

  # main tryCatch to ensure consistent return on error
  tryCatch({

    # 1) Predictor selection (unchanged: assuming you have results_table_numeric / anova available)
    selected_numeric <- results_table_numeric$Variable[results_table_numeric$P_value < alpha_cor]
    selected_factors <- results_table_anova$Variable[results_table_anova$P_value < alpha_anova]
    selected_predictors <- c(selected_numeric, selected_factors)

    if (length(selected_predictors) == 0) {
      return(list(metrics_table = tibble::tibble(
        alpha_cor = alpha_cor, alpha_anova = alpha_anova, Method = c("missForest_two_stage", "xgboost"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_
      ), all_predictions = NULL))
    }

    # 2) Restrict to rows with predictors present (complete cases on predictors)
    predictor_complete_idx <- complete.cases(df[, selected_predictors, drop = FALSE])
    model_df <- df[predictor_complete_idx, c(target_var, selected_predictors), drop = FALSE]

    # 3) Known indices for target
    known_idx <- which(!is.na(model_df[[target_var]]))
    if (length(known_idx) < (min_rows_train + 2)) {
      return(list(metrics_table = tibble::tibble(
        alpha_cor = alpha_cor, alpha_anova = alpha_anova, Method = c("missForest_two_stage", "xgboost"),
        RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_
      ), all_predictions = NULL))
    }

    # 4) Mask selection (keep same logic as before)
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    mask_indices_final <- NULL

    valid_mask <- function(mask_idx, df_model, factor_predictors) {
      train_idx <- setdiff(which(!is.na(df_model[[target_var]])), mask_idx)
      if (length(train_idx) < min_rows_train) return(FALSE)
      for (f in factor_predictors) {
        if (!is.factor(df_model[[f]])) next
        nlev <- nlevels(droplevels(df_model[[f]][train_idx]))
        if (nlev < 2) return(FALSE)
      }
      TRUE
    }
    factor_preds_in_selected <- selected_factors[selected_factors %in% selected_predictors]

    while (attempt <= max_mask_attempts) {
      mask_idx_try <- sample(known_idx, mask_n)
      if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
        mask_indices_final <- mask_idx_try
        break
      }
      attempt <- attempt + 1
    }
    if (is.null(mask_indices_final)) {
      found <- FALSE
      for (mf_try in seq(mask_frac/2, 0.02, by = -0.01)) {
        mask_n2 <- max(1, round(length(known_idx) * mf_try))
        attempt2 <- 1
        while (attempt2 <= max_mask_attempts) {
          mask_idx_try <- sample(known_idx, mask_n2)
          if (valid_mask(mask_idx_try, model_df, factor_preds_in_selected)) {
            mask_indices_final <- mask_idx_try
            found <- TRUE
            break
          }
          attempt2 <- attempt2 + 1
        }
        if (found) break
      }
    }
    if (is.null(mask_indices_final)) stop("Unable to find a valid mask that preserves factor levels")

    # 5) Prepare masked df, BUT per option B we will collapse rare levels only AFTER masking
    df_model_masked <- model_df
    true_values_orig <- df_model_masked[[target_var]][mask_indices_final]
    df_model_masked[[target_var]][mask_indices_final] <- NA
    train_idx <- which(!is.na(df_model_masked[[target_var]]))
    test_idx  <- mask_indices_final

    # 6) Now collapse rare factor levels on the masked subset (both train + test together)
    factor_vars_here <- intersect(selected_factors, names(df_model_masked))
    if (length(factor_vars_here) > 0) {
      df_model_masked <- collapse_rare_levels(df_model_masked, factor_vars_here, min_obs_per_level)
    }

    # 7) Drop single-level factors / constant numerics after collapsing
    final_predictors <- selected_predictors
    if (length(selected_factors) > 0) {
      for (f in selected_factors) {
        if (f %in% final_predictors) {
          nlev_train <- nlevels(droplevels(df_model_masked[[f]][train_idx]))
          if (nlev_train < 2) final_predictors <- setdiff(final_predictors, f)
        }
      }
    }
    for (nvar in selected_numeric) {
      if (nvar %in% final_predictors) {
        if (var(df_model_masked[[nvar]][train_idx], na.rm = TRUE) == 0) final_predictors <- setdiff(final_predictors, nvar)
      }
    }
    if (length(final_predictors) == 0) stop("No valid predictors remain for modeling after collapsing / dropping constants")

    # 8) Build train/test frames with final_predictors
    train_df_reg <- df_model_masked[train_idx, c(target_var, final_predictors), drop = FALSE]
    test_df_reg  <- df_model_masked[test_idx,  c(target_var, final_predictors), drop = FALSE]

    # align factor levels
    for (f in final_predictors) {
      if (is.factor(train_df_reg[[f]])) {
        levs <- levels(droplevels(train_df_reg[[f]]))
        train_df_reg[[f]] <- factor(train_df_reg[[f]], levels = levs)
        test_df_reg[[f]]  <- factor(test_df_reg[[f]], levels = levs)
      }
    }

    # 9) Create LF_log target column in train; test LF_log is NA
    train_df_reg$LF_log <- log1p(train_df_reg[[target_var]])
    test_df_reg$LF_log <- NA_real_

    # helper to build combined matrix for missForest (train + test)
    build_combined_for_impute <- function(train_block, test_block, keep_cols) {
      # select keep_cols present
      keep_cols_present <- keep_cols[keep_cols %in% names(train_block)]
      train_sel <- train_block[, keep_cols_present, drop = FALSE]
      test_sel  <- test_block[, keep_cols_present, drop = FALSE]
      safe_rbind(train_sel, test_sel)
    }

    # 10) CROSS-VALIDATION on training set to get CV predictions (for residuals)
    ntrain <- nrow(train_df_reg)
    folds <- sample(rep(1:cv_folds, length.out = ntrain))
    cv_pred_train_log <- rep(NA_real_, ntrain)

    # columns to include in imputation: final_predictors + LF_log (LF_log exists in train)
    impute_cols <- c(final_predictors, "LF_log")

    for (k in seq_len(cv_folds)) {
      tr_idx_k <- which(folds != k)
      val_idx_k <- which(folds == k)
      if (length(tr_idx_k) < 5 || length(val_idx_k) < 1) next

      fold_train <- train_df_reg[tr_idx_k, , drop = FALSE]
      fold_val   <- train_df_reg[val_idx_k, , drop = FALSE]

      combined_fold <- build_combined_for_impute(fold_train[, impute_cols, drop = FALSE],
                                                fold_val[, impute_cols, drop = FALSE],
                                                impute_cols)
      # impute LF_log on combined_fold using missForest
      mf_out <- tryCatch(missForest::missForest(combined_fold, maxiter = 5, ntree = mf_ntree, verbose = FALSE),
                         error = function(e) NULL)
      if (is.null(mf_out)) next

      imputed_fold <- mf_out$ximp
      ntr <- nrow(fold_train)
      pred_val_log <- imputed_fold$LF_log[(ntr + 1):nrow(imputed_fold)]
      cv_pred_train_log[val_idx_k] <- pred_val_log
    }

    # if some cv preds missing, fill them by fitting on full train once
    missing_cv <- which(is.na(cv_pred_train_log))
    if (length(missing_cv) > 0) {
      combined_full <- build_combined_for_impute(train_df_reg[, impute_cols, drop = FALSE],
                                                test_df_reg[, impute_cols, drop = FALSE],
                                                impute_cols)
      mf_out_full <- tryCatch(missForest::missForest(combined_full, maxiter = 5, ntree = mf_ntree, verbose = FALSE),
                              error = function(e) NULL)
      if (!is.null(mf_out_full)) {
        imputed_full <- mf_out_full$ximp
        cv_train_full <- imputed_full$LF_log[1:nrow(train_df_reg)]
        cv_pred_train_log[missing_cv] <- cv_train_full[missing_cv]
      }
    }

    # 11) compute residuals on original scale and determine hard training rows
    pred_train_orig <- expm1(cv_pred_train_log)
    true_train_orig <- train_df_reg[[target_var]]
    resid_train <- abs(true_train_orig - pred_train_orig)
    resid_mean <- mean(resid_train, na.rm = TRUE)
    resid_sd   <- sd(resid_train, na.rm = TRUE)
    hard_threshold <- resid_mean + hard_resid_sd_mult * resid_sd
    hard_train_idx <- which(resid_train > hard_threshold)

    # relax threshold until we have at least specialist_min_rows (conservative)
    if (length(hard_train_idx) < specialist_min_rows) {
      mult_try <- hard_resid_sd_mult
      while (length(hard_train_idx) < specialist_min_rows && mult_try >= 0.5) {
        mult_try <- mult_try - 0.25
        hard_threshold2 <- resid_mean + mult_try * resid_sd
        hard_train_idx <- which(resid_train > hard_threshold2)
        if (mult_try <= 0.5) break
      }
    }

    # 12) MAIN imputation using missForest (train+test combined)
    combined_main <- build_combined_for_impute(train_df_reg[, impute_cols, drop = FALSE],
                                              test_df_reg[, impute_cols, drop = FALSE],
                                              impute_cols)
    mf_out_main <- tryCatch(missForest::missForest(combined_main, maxiter = 10, ntree = mf_ntree, verbose = FALSE),
                            error = function(e) NULL)
    main_pred_log <- rep(NA_real_, nrow(test_df_reg))
    if (!is.null(mf_out_main)) {
      imputed_main <- mf_out_main$ximp
      ntr <- nrow(train_df_reg)
      main_pred_log <- imputed_main$LF_log[(ntr + 1):nrow(imputed_main)]
    }

    # 13) SPECIALIST imputation (if enough hard rows)
    specialist_pred_log <- rep(NA_real_, nrow(test_df_reg))
    if (length(hard_train_idx) >= 5) {
      spec_train <- train_df_reg[hard_train_idx, , drop = FALSE]
      spec_impute_cols <- c(final_predictors, "LF_log")
      spec_combined <- build_combined_for_impute(spec_train[, spec_impute_cols, drop = FALSE],
                                                 test_df_reg[, spec_impute_cols, drop = FALSE],
                                                 spec_impute_cols)
      mf_out_spec <- tryCatch(missForest::missForest(spec_combined, maxiter = 10, ntree = max(200, mf_ntree), verbose = FALSE),
                              error = function(e) NULL)
      if (!is.null(mf_out_spec)) {
        imputed_spec <- mf_out_spec$ximp
        ntr_spec <- nrow(spec_train)
        specialist_pred_log <- imputed_spec$LF_log[(ntr_spec + 1):nrow(imputed_spec)]
      }

      # classifier to route test rows to specialist (random forest)
      is_hard_label <- rep(0, nrow(train_df_reg)); is_hard_label[hard_train_idx] <- 1
      clf_df <- train_df_reg[, final_predictors, drop = FALSE]
      clf_df$is_hard <- factor(is_hard_label)
      # remove zero-variance columns
      keep_cols <- names(clf_df)[sapply(clf_df, function(x) !(is.numeric(x) && var(as.numeric(x), na.rm = TRUE) == 0))]
      clf_df <- clf_df[, keep_cols, drop = FALSE]
      rf_clf <- tryCatch(randomForest::randomForest(is_hard ~ ., data = clf_df, ntree = 200), error = function(e) NULL)
      use_specialist <- integer(0)
      if (!is.null(rf_clf)) {
        test_clf_df <- test_df_reg[, names(clf_df)[names(clf_df) != "is_hard"], drop = FALSE]
        # align factor levels with training clf_df
        for (f in names(test_clf_df)) if (is.factor(clf_df[[f]])) test_clf_df[[f]] <- factor(test_clf_df[[f]], levels = levels(clf_df[[f]]))
        probs <- tryCatch(predict(rf_clf, newdata = test_clf_df, type = "prob")[, "1"], error = function(e) rep(0, nrow(test_clf_df)))
        use_specialist <- which(probs > specialist_prob_threshold & !is.na(specialist_pred_log))
      }
    }

    # 14) Combine main & specialist predictions, back-transform
    final_pred_log <- main_pred_log
    if (length(use_specialist) > 0) final_pred_log[use_specialist] <- specialist_pred_log[use_specialist]
    final_pred_orig_mf <- expm1(final_pred_log)

    # 15) XGBoost model (single-stage) - train on train_df_reg, predict test_df_reg
    # Prepare model.matrix (one-hot encoding) for numeric + factor predictors
    # Build formula to include only final_predictors
    xgb_train_mat <- tryCatch({
      model.matrix(~ . - 1, data = train_df_reg[, final_predictors, drop = FALSE])
    }, error = function(e) NULL)
    xgb_test_mat <- NULL
    xgb_pred_orig <- rep(NA_real_, nrow(test_df_reg))
    if (!is.null(xgb_train_mat)) {
      # ensure same columns/order for test
      xgb_test_mat <- tryCatch({
        model.matrix(~ . - 1, data = test_df_reg[, final_predictors, drop = FALSE])
      }, error = function(e) NULL)
      if (!is.null(xgb_test_mat)) {
        # align columns
        allcols <- union(colnames(xgb_train_mat), colnames(xgb_test_mat))
        add_missing_mat <- function(mat, cols) {
          for (cname in cols) if (!(cname %in% colnames(mat))) mat <- cbind(mat, setNames(rep(0, nrow(mat)), cname))
          mat[, cols, drop = FALSE]
        }
        xgb_train_mat2 <- add_missing_mat(xgb_train_mat, allcols)
        xgb_test_mat2  <- add_missing_mat(xgb_test_mat, allcols)
        # order columns
        xgb_train_mat2 <- xgb_train_mat2[, allcols, drop = FALSE]
        xgb_test_mat2  <- xgb_test_mat2[, allcols, drop = FALSE]
        # target is LF_log
        y_train_log <- train_df_reg$LF_log
        dtrain <- xgboost::xgb.DMatrix(data = xgb_train_mat2, label = y_train_log)
        params <- list(objective = "reg:squarederror", eval_metric = "rmse")
        xgb_mod <- tryCatch(xgboost::xgb.train(params = params, data = dtrain, nrounds = xgb_nrounds, verbose = 0), error = function(e) NULL)
        if (!is.null(xgb_mod)) {
          dtest <- xgboost::xgb.DMatrix(data = xgb_test_mat2)
          pred_xgb_log <- tryCatch(as.numeric(predict(xgb_mod, dtest)), error = function(e) rep(NA_real_, nrow(test_df_reg)))
          xgb_pred_orig <- expm1(pred_xgb_log)
        }
      }
    }

    # 16) Compose predictions tibble and compute metrics
    all_predictions <- tibble::tibble(
      index = test_idx,
      true = true_values_orig,
      pred_missForest_two_stage = final_pred_orig_mf,
      pred_xgboost = xgb_pred_orig
    )

    res_mf <- eval_metrics(true_values_orig, all_predictions$pred_missForest_two_stage)
    res_xgb <- eval_metrics(true_values_orig, all_predictions$pred_xgboost)

    metrics_table <- tibble::tibble(
      alpha_cor = alpha_cor,
      alpha_anova = alpha_anova,
      mf_ntree = mf_ntree,
      xgb_nrounds = xgb_nrounds,
      Method = c("missForest_two_stage", "xgboost"),
      RMSE = c(res_mf$RMSE, res_xgb$RMSE),
      MAE  = c(res_mf$MAE, res_xgb$MAE),
      R2   = c(res_mf$R2, res_xgb$R2),
      n_predictors = length(final_predictors)
    )

    # return requested outputs: best prediction per method? we return per-method predictions + metrics
    return(list(
      metrics_table = metrics_table,
      all_predictions = all_predictions
    ))

  }, error = function(e) {
    warning("evaluate_combo failed: ", conditionMessage(e))
    tib <- tibble::tibble(
      alpha_cor = if (exists("alpha_cor")) alpha_cor else NA_real_,
      alpha_anova = if (exists("alpha_anova")) alpha_anova else NA_real_,
      mf_ntree = if (exists("mf_ntree")) mf_ntree else NA_integer_,
      xgb_nrounds = if (exists("xgb_nrounds")) xgb_nrounds else NA_integer_,
      Method = c("missForest_two_stage", "xgboost"),
      RMSE = NA_real_, MAE = NA_real_, R2 = NA_real_, n_predictors = NA_integer_
    )
    return(list(metrics_table = tib, all_predictions = NULL))
  })
}

results_list <- purrr::map(param_list, evaluate_combo, df = train_csv)
# collect metrics:
metrics_all <- purrr::map_dfr(results_list, ~ .x$metrics_table)
# collect predictions (list per combo):
preds_all <- purrr::map(results_list, ~ .x$all_predictions)

best_by_combo <- metrics_all %>%
  group_by(alpha_cor, alpha_anova, mf_ntree, xgb_nrounds, Method) %>%
  summarise(RMSE = mean(RMSE, na.rm = TRUE), .groups = "drop") %>%
  arrange(RMSE) %>%
  slice_head(n = 20)

print(best_by_combo)
```
This updated method produces a very similar result, (R2 = 0.69, RMSe = 11.4), although not a better one. Also xgboost works really well, almost as well as missForest on this dataset. 

We'll next try to see if the model can be fit on the errors present in the inaccurate predictions. If the errors are representative of some underlying pattern, this can improve the predictions by fitting for these specific patterns. However, this method is prone to some level of overfitting, so we'll check using cross-validation to ensure that any improvement we see in prediction accuracy is due to the updated model fitting patterns in the errors and not noise within the error terms.
```{r}
refine_imputation_cv_rf <- function(df_all,
                                    best_prediction,
                                    predictors,
                                    target_var = "LotFrontage",
                                    k = 10,
                                    ntree = 100,
                                    nodesize = 5,
                                    numeric_top_n = 5,
                                    seed = 123,
                                    verbose = TRUE) {
  # Dependencies
  suppressPackageStartupMessages({
    library(dplyr)
    library(randomForest)
    library(ggplot2)
  })
  set.seed(seed)

  # --- Sanity checks -------------------------------------------------------
  required_cols <- c("pred", "true", "index")
  if (!all(required_cols %in% names(best_prediction))) {
    stop("best_prediction must contain columns: 'pred', 'true', 'index'.")
  }
  if (!all(best_prediction$index %in% seq_len(nrow(df_all)))) {
    stop("Some indices in best_prediction$index are not valid row indices in df_all.")
  }

  # Observed indices and vectors (aligned to best_prediction order)
  obs_idx <- best_prediction$index
  preds_obs <- best_prediction$pred
  true_obs  <- best_prediction$true
  n_obs <- length(obs_idx)
  if (n_obs < k) stop("Number of folds k must be <= number of observed rows.")

  # Prepare full-length vectors (aligned to df_all) for diagnostic function
  n_all <- nrow(df_all)
  preds_full <- rep(NA_real_, n_all)
  true_full  <- rep(NA_real_, n_all)
  preds_full[obs_idx] <- preds_obs
  true_full[obs_idx]  <- true_obs

  # Run your diagnostic function (expects full-length vectors)
  diagnostic <- analyse_bad_predictions(
    df_all = df_all,
    predictions = preds_full,
    true_values = true_full,
    predictors = predictors,
    n_sd = 0.5,
    p_thresh = 0.1
  )

  # --- Choose problem variables (from diagnostic) -------------------------
  problem_vars <- character(0)

  if (!is.null(diagnostic$numeric_summary) && nrow(diagnostic$numeric_summary) > 0) {
    ordered_nums <- diagnostic$numeric_summary %>%
      arrange(desc(abs(Mean_Diff))) %>%
      pull(Variable)
    problem_vars <- c(problem_vars, head(ordered_nums, numeric_top_n))
  }
  if (!is.null(diagnostic$factor_summary) && length(diagnostic$factor_summary) > 0) {
    problem_vars <- c(problem_vars, names(diagnostic$factor_summary))
  }
  problem_vars <- unique(problem_vars)
  # Fallback to the full predictors if diagnostic gave nothing usable
  if (length(problem_vars) == 0) {
    if (verbose) message("Diagnostic did not flag variables — using all predictors.")
    problem_vars <- predictors
  } else {
    # ensure problem_vars exist in predictors
    problem_vars <- intersect(problem_vars, predictors)
    if (length(problem_vars) == 0) {
      if (verbose) message("Problem vars were not found in provided predictors — using all predictors.")
      problem_vars <- predictors
    }
  }

  # --- Build modeling dataframe for observed rows --------------------------
  df_obs <- df_all[obs_idx, problem_vars, drop = FALSE] %>% as.data.frame()
  df_obs$.residual <- preds_obs - true_obs  # r = pred - true

  # --- k-fold cross-validation (OOF predictions for the residuals) -------
  folds <- sample(rep(seq_len(k), length.out = n_obs))
  oof_pred_residual <- rep(NA_real_, n_obs)   # predicted residuals out-of-fold
  oof_pred_corrected <- rep(NA_real_, n_obs)  # corrected predictions out-of-fold

  for (fold in seq_len(k)) {
    train_idx <- which(folds != fold)
    val_idx   <- which(folds == fold)

    train_x <- df_obs[train_idx, problem_vars, drop = FALSE]
    train_y <- df_obs$.residual[train_idx]

    val_x <- df_obs[val_idx, problem_vars, drop = FALSE]

    # Train a restrained random forest on residuals
    rf_model <- randomForest(x = train_x, y = train_y, ntree = ntree, nodesize = nodesize)

    # Predict residuals on validation fold
    pred_res_val <- predict(rf_model, val_x)

    # store OOF predictions
    oof_pred_residual[val_idx] <- pred_res_val
    # corrected predictions on validation fold (in best_prediction order)
    oof_pred_corrected[val_idx] <- preds_obs[val_idx] - pred_res_val
  }

  # --- CV metrics (on observed rows, using OOF predictions) ----------------
  rmse_before <- sqrt(mean((preds_obs - true_obs)^2, na.rm = TRUE))
  rmse_after_oof <- sqrt(mean((oof_pred_corrected - true_obs)^2, na.rm = TRUE))
  improvement_oof <- rmse_before - rmse_after_oof

  if (verbose) {
    cat(sprintf("OOF CV: RMSE before = %.4f; after (OOF corrected) = %.4f; Δ = %.4f\n",
                rmse_before, rmse_after_oof, improvement_oof))
  }

  # --- Train final meta-model on all observed rows -------------------------
  final_rf <- randomForest(x = df_obs[, problem_vars, drop = FALSE],
                           y = df_obs$.residual,
                           ntree = ntree,
                           nodesize = nodesize)

  # Predict correction for all rows in df_all (will return NA for rows where model.matrix fails,
  # but randomForest supports factors so it's fine if df_all has same factor levels).
  df_all_sub <- df_all[, problem_vars, drop = FALSE] %>% as.data.frame()
  final_correction_full <- tryCatch({
    predict(final_rf, newdata = df_all_sub)
  }, error = function(e) {
    # if predict fails (rare), produce NA vector
    warning("Predicting final correction on full data failed: ", e$message)
    rep(NA_real_, n_all)
  })

  # Build corrected predictions
  corrected_full <- preds_full - final_correction_full   # only where preds_full not NA will be meaningful
  corrected_obs <- corrected_full[obs_idx]              # same order as best_prediction

  # --- More summary metrics -----------------------------------------------
  rmse_after_final <- sqrt(mean((corrected_obs - true_obs)^2, na.rm = TRUE))
  improvement_final <- rmse_before - rmse_after_final

  if (verbose) {
    cat(sprintf("Final model trained on all observed rows: RMSE after final correction = %.4f (Δ relative to before = %.4f)\n",
                rmse_after_final, improvement_final))
  }

  # --- Generate diagnostic plots ------------------------------------------
  # 1) Residual histograms before / after (OOF)
  plot_resid_df <- data.frame(
    type = rep(c("before", "after_oof"), each = n_obs),
    resid = c((preds_obs - true_obs), (oof_pred_residual - 0)) # note: before residual = pred - true; after_oof residual ~ corrected - true = (pred - pred_resid) - true = (pred - true) - pred_resid
  )
  # But easier: compare distribution of (pred - true) and (corrected - true)
  plot_resid_df2 <- data.frame(
    type = rep(c("before", "after_oof"), each = n_obs),
    value = c((preds_obs - true_obs), (oof_pred_corrected - true_obs))
  )

  p1 <- ggplot(plot_resid_df2, aes(x = value, fill = type)) +
    geom_density(alpha = 0.35) + labs(title = "Residual distribution: before vs OOF-corrected",
                                      x = "Residual (pred - true)", y = "Density") +
    theme_minimal()

  # 2) Scatter: predicted vs true (before and after OOF) — sample if many points
  plot_df_scatter <- data.frame(
    true = rep(true_obs, 2),
    pred = c(preds_obs, oof_pred_corrected),
    stage = rep(c("before", "after_oof"), each = n_obs)
  )
  p2 <- ggplot(plot_df_scatter, aes(x = true, y = pred, color = stage)) +
    geom_point(alpha = 0.7) +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
    labs(title = "Predicted vs True: before vs OOF-corrected", x = "True", y = "Predicted") +
    theme_minimal()

  # 3) Improvement by top factor (if any factor vars exist among problem_vars)
  factor_vars_in_problem <- problem_vars[sapply(df_all[, problem_vars, drop = FALSE], is.factor) |
                                          sapply(df_all[, problem_vars, drop = FALSE], is.character)]
  p3 <- NULL
  if (length(factor_vars_in_problem) > 0) {
    top_factor <- factor_vars_in_problem[1]
    df_factor <- data.frame(
      level = as.character(df_all[obs_idx, top_factor]),
      before_err = preds_obs - true_obs,
      after_err_oof = oof_pred_corrected - true_obs
    )
    agg_df <- df_factor %>%
      group_by(level) %>%
      summarise(mean_before = mean(before_err, na.rm = TRUE),
                mean_after  = mean(after_err_oof, na.rm = TRUE),
                n = n()) %>%
      arrange(desc(abs(mean_before))) %>%
      filter(n >= 3)  # at least 3 observations per level for stable display

    if (nrow(agg_df) > 0) {
      agg_df_long <- tidyr::pivot_longer(agg_df, cols = c("mean_before", "mean_after"),
                                         names_to = "stage", values_to = "mean_err")
      p3 <- ggplot(agg_df_long, aes(x = reorder(level, -abs(mean_err)), y = mean_err, fill = stage)) +
        geom_col(position = position_dodge(width = 0.8)) +
        labs(title = paste0("Mean residual by level of ", top_factor),
             x = paste0(top_factor, " (levels)"), y = "Mean residual (pred - true)") +
        theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
    }
  }

  # --- Return a structured list -------------------------------------------
  out <- list(
    diagnostic = diagnostic,         # from analyse_bad_predictions (full-df aligned)
    problem_vars = problem_vars,     # predictors used by meta-model
    final_model = final_rf,          # final randomForest meta-model (trained on all observed rows)
    oof = list(
      residual_pred_oof = oof_pred_residual,   # predicted residuals, OOF, length n_obs
      corrected_pred_oof = oof_pred_corrected, # corrected predictions, OOF, length n_obs
      rmse_before = rmse_before,
      rmse_after_oof = rmse_after_oof,
      improvement_oof = improvement_oof
    ),
    final = list(
      corrected_full = corrected_full, # length n_all; NA where preds_full is NA
      corrected_obs = corrected_obs,   # length n_obs; same order as best_prediction
      rmse_after_final = rmse_after_final,
      improvement_final = improvement_final
    ),
    plots = list(resid_density = p1, pred_vs_true = p2, factor_improvement = p3),
    meta = list(k = k, ntree = ntree, nodesize = nodesize, seed = seed,
                problem_var_count = length(problem_vars))
  )

  return(out)
}


res <- refine_imputation_cv_rf(
  df_all = df_all,
  best_prediction = best_prediction,   # must have columns 'pred', 'true', 'index'
  predictors = names(df_all),          # or a subset of predictors
  target_var = "LotFrontage",
  k = 10,
  ntree = 100,
  nodesize = 5,
  numeric_top_n = 5,
  seed = 42,
  verbose = TRUE
)
```

```{r}
fe_engineering <- function(df,
                           collapse_factor_min = 5,
                           numeric_log = NULL,
                           numeric_sqrt = NULL,
                           poly_vars = NULL,        # named list: list(col = degree)
                           interactions = NULL,     # list of character vectors c("A","B")
                           binning = NULL,          # named list: list(col = c(cuts))
                           winsorize = NULL,        # named list: list(col = c(lower_pct, upper_pct))
                           one_hot = FALSE,
                           remove_original = FALSE) {
  # Purpose: create common FE transforms in-place and return metadata
  #
  # Returns: list(data = data.frame, meta = list(created = <vector>, mapping info...))
  #
  library(dplyr); library(forcats); library(tidyr)
  df2 <- as.data.frame(df)
  meta <- list(created = character(0), collapsed = list())

  # 1) collapse rare factor levels
  factor_cols <- names(df2)[sapply(df2, is.factor)]
  for (v in factor_cols) {
    tbl <- table(df2[[v]])
    rare <- names(tbl)[tbl < collapse_factor_min]
    if (length(rare) > 0) {
      df2[[v]] <- forcats::fct_collapse(df2[[v]], Other = rare)
      df2[[v]] <- droplevels(df2[[v]])
      meta$collapsed[[v]] <- rare
    }
  }

  # 2) winsorize numeric columns
  if (!is.null(winsorize)) {
    for (v in names(winsorize)) {
      p <- winsorize[[v]]
      if (!is.numeric(df2[[v]])) next
      lo <- quantile(df2[[v]], probs = p[1], na.rm = TRUE)
      hi <- quantile(df2[[v]], probs = p[2], na.rm = TRUE)
      newn <- paste0(v, "_winsor")
      df2[[newn]] <- pmin(pmax(df2[[v]], lo), hi)
      meta$created <- c(meta$created, newn)
    }
  }

  # 3) log / sqrt transforms
  if (!is.null(numeric_log)) {
    for (v in numeric_log) {
      if (!is.numeric(df2[[v]])) next
      newn <- paste0("log_", v)
      df2[[newn]] <- ifelse(!is.na(df2[[v]]) & df2[[v]] > 0, log(df2[[v]]), NA_real_)
      meta$created <- c(meta$created, newn)
    }
  }
  if (!is.null(numeric_sqrt)) {
    for (v in numeric_sqrt) {
      if (!is.numeric(df2[[v]])) next
      newn <- paste0("sqrt_", v)
      df2[[newn]] <- ifelse(!is.na(df2[[v]]), sqrt(pmax(df2[[v]], 0)), NA_real_)
      meta$created <- c(meta$created, newn)
    }
  }

  # 4) polynomial features
  if (!is.null(poly_vars)) {
    for (v in names(poly_vars)) {
      deg <- as.integer(poly_vars[[v]])
      if (!is.numeric(df2[[v]])) next
      for (d in 2:deg) {
        newn <- paste0(v, "_poly", d)
        df2[[newn]] <- ifelse(!is.na(df2[[v]]), df2[[v]]^d, NA_real_)
        meta$created <- c(meta$created, newn)
      }
    }
  }

  # 5) interactions (simple multiplicative interactions)
  if (!is.null(interactions)) {
    for (pair in interactions) {
      v1 <- pair[1]; v2 <- pair[2]
      if (! (v1 %in% names(df2) && v2 %in% names(df2))) next
      newn <- paste0(v1, "_x_", v2)
      df2[[newn]] <- ifelse(!is.na(df2[[v1]]) & !is.na(df2[[v2]]), df2[[v1]] * df2[[v2]], NA_real_)
      meta$created <- c(meta$created, newn)
    }
  }

  # 6) binning
  if (!is.null(binning)) {
    for (v in names(binning)) {
      if (!is.numeric(df2[[v]])) next
      cuts <- binning[[v]]
      newn <- paste0(v, "_binned")
      df2[[newn]] <- cut(df2[[v]], breaks = c(-Inf, cuts, Inf), include.lowest = TRUE)
      df2[[newn]] <- droplevels(df2[[newn]])
      meta$created <- c(meta$created, newn)
    }
  }

  # 7) optional one-hot encoding
  df_out <- df2
  if (one_hot) {
    mm <- model.matrix(~ . - 1, data = df2)
    df_out <- as.data.frame(mm)
  }

  # 8) remove originals if requested
  if (remove_original) {
    to_remove <- c(names(poly_vars), numeric_log, numeric_sqrt, names(binning), unlist(interactions))
    to_remove <- intersect(to_remove, names(df_out))
    df_out <- df_out[, setdiff(names(df_out), to_remove), drop = FALSE]
  }

  return(list(data = df_out, meta = meta))
}


```

```{r}
run_imputer_cv <- function(df,
                           target_var = "LotFrontage",
                           predictors = NULL,
                           methods = c("lm", "xgb", "lgb", "ranger"),
                           param_grids = NULL,         # list of param grids per method (see defaults below)
                           n_folds = 5,
                           n_repeats = 3,
                           mask_frac = 0.20,           # constant 20% masked per fold as requested
                           min_rows_train = 30,
                           seed = 123,
                           parallel = TRUE,
                           workers = NULL,
                           verbose = TRUE) {
  library(dplyr); library(purrr); library(tidyr)
  library(furrr); library(future)
  library(ranger); library(randomForest)
  library(xgboost); library(lightgbm)
  library(mgcv); # for GAM if you ever add gam to methods
  set.seed(seed)

  if (is.null(predictors)) predictors <- setdiff(names(df), target_var)
  if (!target_var %in% names(df)) stop("target_var not in df")

  # prepare default param grids if none supplied
  default_grids <- list(
    lm = expand.grid(dummy = NA),  # lm has no hyperparams
    xgb = expand.grid(eta = c(0.05, 0.1),
                      max_depth = c(3L, 5L),
                      nrounds = c(50L, 100L)),
    lgb = expand.grid(learning_rate = c(0.05, 0.1),
                      num_leaves = c(15L, 31L),
                      nrounds = c(50L, 100L)),
    ranger = expand.grid(num.trees = c(100L),
                         min.node.size = c(5L, 10L))
  )
  if (is.null(param_grids)) param_grids <- default_grids
  # ensure each requested method has a param grid
  for (m in methods) {
    if (!m %in% names(param_grids)) {
      if (m %in% names(default_grids)) param_grids[[m]] <- default_grids[[m]] else stop("No param grid for method ", m)
    }
  }

  # setup parallel plan
  if (parallel) {
    if (is.null(workers)) {
      workers <- max(1L, future::availableCores() - 1L)
    }
    plan(multisession, workers = workers)
    if (verbose) message("Parallel plan multisession with ", workers, " workers")
    use_future_map <- TRUE
  } else {
    plan(sequential)
    use_future_map <- FALSE
  }

  # helper: mask selection that preserves factor levels in train
  make_mask_indices <- function(known_idx, mask_frac, df_local, predictors_local, min_rows_train, max_attempts = 200) {
    mask_n <- max(1, round(length(known_idx) * mask_frac))
    attempt <- 1
    while (attempt <= max_attempts) {
      mask_try <- sample(known_idx, mask_n)
      train_idx <- setdiff(known_idx, mask_try)
      if (length(train_idx) < min_rows_train) { attempt <- attempt + 1; next }
      # ensure factors have >1 level in training
      ok <- TRUE
      for (f in predictors_local[sapply(df_local[, predictors_local, drop = FALSE], is.factor)]) {
        if (nlevels(droplevels(df_local[[f]][train_idx])) < 2) { ok <- FALSE; break }
      }
      if (ok) return(list(mask = mask_try, train = train_idx))
      attempt <- attempt + 1
    }
    return(NULL)
  }

  # get indices where target is observed
  obs_all_idx <- which(!is.na(df[[target_var]]))
  if (length(obs_all_idx) < (min_rows_train + 2)) stop("Not enough observed target rows to perform CV")

  # For results storage
  all_method_results <- list()

  # Iterate methods
  for (method in methods) {
    if (verbose) message("Evaluating method: ", method)
    grid <- param_grids[[method]]
    # make grid list of rows
    grid_list <- split(grid, seq_len(nrow(grid)))

    # evaluate each param combo with repeated masking CV
    eval_param_combo <- function(param_row) {
      params <- as.list(param_row)
      if (verbose) message("  combo: ", paste(names(params), params, collapse = ", "))
      # storage for OOF predictions and truths
      oof_preds <- rep(NA_real_, length(obs_all_idx))
      names(oof_preds) <- obs_all_idx
      truths <- df[[target_var]][obs_all_idx]
      # repeated masking
      for (rep_i in seq_len(n_repeats)) {
        for (fold_i in seq_len(n_folds)) {
          mask_out <- make_mask_indices(obs_all_idx, mask_frac = mask_frac, df_local = df, predictors_local = predictors, min_rows_train = min_rows_train)
          if (is.null(mask_out)) next
          mask_idx <- mask_out$mask
          train_idx <- mask_out$train

          # Build training dataframe (complete-case for chosen predictors)
          keep_idx <- c(train_idx, mask_idx)
          model_df <- df[keep_idx, c(target_var, predictors), drop = FALSE]
          # set masked values to NA
          model_df[[target_var]][match(mask_idx, keep_idx)] <- NA

          train_rows <- which(!is.na(model_df[[target_var]]))
          test_rows_rel <- which(is.na(model_df[[target_var]])) # relative to model_df
          if (length(train_rows) < min_rows_train) next

          # align factor levels for predictors within model_df
          for (p in predictors) {
            if (is.factor(model_df[[p]])) {
              levs <- levels(droplevels(model_df[[p]][train_rows]))
              model_df[[p]] <- factor(model_df[[p]], levels = levs)
            }
          }

          # prepare data for each method
          if (method == "lm") {
            form <- as.formula(paste(target_var, "~", paste(predictors, collapse = " + ")))
            fit <- tryCatch(lm(form, data = model_df[train_rows, , drop = FALSE]), error = function(e) NULL)
            if (!is.null(fit)) {
              preds_rel <- tryCatch(predict(fit, newdata = model_df[test_rows_rel, , drop = FALSE]), error = function(e) rep(NA_real_, length(test_rows_rel)))
            } else preds_rel <- rep(NA_real_, length(test_rows_rel))
          } else if (method == "xgb") {
            mm_train <- model.matrix(~ . - 1, data = model_df[train_rows, predictors, drop = FALSE])
            mm_test  <- model.matrix(~ . - 1, data = model_df[test_rows_rel, predictors, drop = FALSE])
            dtrain <- xgboost::xgb.DMatrix(data = mm_train, label = model_df[[target_var]][train_rows])
            xgb_params <- list(objective = "reg:squarederror", eta = params$eta, max_depth = as.integer(params$max_depth))
            xgb_mod <- tryCatch(xgboost::xgb.train(params = xgb_params, data = dtrain, nrounds = as.integer(params$nrounds), verbose = 0), error = function(e) NULL)
            if (!is.null(xgb_mod)) {
              preds_rel <- tryCatch(predict(xgb_mod, mm_test), error = function(e) rep(NA_real_, nrow(mm_test)))
            } else preds_rel <- rep(NA_real_, nrow(mm_test))
          } else if (method == "lgb") {
            mm_train <- model.matrix(~ . - 1, data = model_df[train_rows, predictors, drop = FALSE])
            mm_test  <- model.matrix(~ . - 1, data = model_df[test_rows_rel, predictors, drop = FALSE])
            dtrain <- lightgbm::lgb.Dataset(data = mm_train, label = model_df[[target_var]][train_rows])
            lgb_params <- list(objective = "regression", learning_rate = params$learning_rate, num_leaves = as.integer(params$num_leaves))
            lgb_mod <- tryCatch(lightgbm::lgb.train(params = lgb_params, data = dtrain, nrounds = as.integer(params$nrounds), verbose = -1), error = function(e) NULL)
            if (!is.null(lgb_mod)) {
              preds_rel <- tryCatch(predict(lgb_mod, mm_test), error = function(e) rep(NA_real_, nrow(mm_test)))
            } else preds_rel <- rep(NA_real_, nrow(mm_test))
          } else if (method == "ranger") {
            # ranger supports factor predictors natively
            fit <- tryCatch(ranger::ranger(as.formula(paste(target_var, "~", paste(predictors, collapse = " + "))),
                                          data = model_df[train_rows, , drop = FALSE],
                                          num.trees = as.integer(params$num.trees),
                                          min.node.size = as.integer(params$min.node.size),
                                          verbose = FALSE), error = function(e) NULL)
            if (!is.null(fit)) {
              preds_rel <- tryCatch(predict(fit, data = model_df[test_rows_rel, , drop = FALSE])$predictions, error = function(e) rep(NA_real_, length(test_rows_rel)))
            } else preds_rel <- rep(NA_real_, length(test_rows_rel))
          } else {
            preds_rel <- rep(NA_real_, length(test_rows_rel))
          }

          # put preds_rel into oof_preds at the absolute indices
          abs_test_idx <- keep_idx[test_rows_rel]
          for (j in seq_along(abs_test_idx)) {
            abs_idx <- abs_test_idx[j]
            oof_preds[as.character(abs_idx)] <- preds_rel[j]
          }

        } # end folds
      } # end repeats

      # compute metrics on oof_preds vs truths (only where both present)
      idx_ok <- which(!is.na(oof_preds))
      if (length(idx_ok) == 0) {
        rmse <- NA_real_; mae <- NA_real_; r2 <- NA_real_ ; n_ok <- 0
      } else {
        tvals <- truths[idx_ok]; pvals <- unname(oof_preds[idx_ok])
        rmse <- sqrt(mean((tvals - pvals)^2))
        mae <- mean(abs(tvals - pvals))
        r2 <- if (length(tvals) > 1) cor(tvals, pvals)^2 else NA_real_
        n_ok <- length(idx_ok)
      }

      list(params = params, rmse = rmse, mae = mae, r2 = r2, n = n_ok, oof_preds = oof_preds)
    } # end eval_param_combo

    # Parallel or serial mapping through grid_list
    if (use_future_map) {
      combos_res <- furrr::future_map(grid_list, eval_param_combo, .progress = verbose)
    } else {
      combos_res <- purrr::map(grid_list, eval_param_combo)
    }

    # pick best combo by rmse (lowest)
    rmses <- sapply(combos_res, function(x) x$rmse)
    best_idx <- which.min(rmses)
    best_combo_res <- combos_res[[best_idx]]

    # Fit final model on all observed rows using best params (to produce final imputations)
    observed_idx <- which(!is.na(df[[target_var]]))
    full_train_df <- df[observed_idx, c(target_var, predictors), drop = FALSE]
    for (p in predictors) {
      if (is.factor(full_train_df[[p]])) full_train_df[[p]] <- droplevels(full_train_df[[p]])
    }

    if (method == "lm") {
      final_fit <- tryCatch(lm(as.formula(paste(target_var, "~", paste(predictors, collapse = " + "))), data = full_train_df), error = function(e) NULL)
      final_preds_full <- rep(NA_real_, nrow(df))
      if (!is.null(final_fit)) {
        rows_to_pred <- which(is.na(df[[target_var]])) # the originally-missing rows
        if (length(rows_to_pred) > 0) {
          newdata <- df[rows_to_pred, predictors, drop = FALSE]
          # align factor levels
          for (p in predictors) if (is.factor(full_train_df[[p]])) newdata[[p]] <- factor(newdata[[p]], levels = levels(full_train_df[[p]]))
          final_preds_full[rows_to_pred] <- tryCatch(predict(final_fit, newdata = newdata), error = function(e) rep(NA_real_, length(rows_to_pred)))
        }
      }
    } else if (method == "xgb") {
      mm_train <- model.matrix(~ . - 1, data = full_train_df[, predictors, drop = FALSE])
      dtrain <- xgboost::xgb.DMatrix(data = mm_train, label = full_train_df[[target_var]])
      bp <- best_combo_res$params
      xgb_params <- list(objective = "reg:squarederror", eta = bp$eta, max_depth = as.integer(bp$max_depth))
      final_fit <- tryCatch(xgboost::xgb.train(params = xgb_params, data = dtrain, nrounds = as.integer(bp$nrounds), verbose = 0), error = function(e) NULL)
      final_preds_full <- rep(NA_real_, nrow(df))
      if (!is.null(final_fit)) {
        rows_to_pred <- which(is.na(df[[target_var]]))
        if (length(rows_to_pred) > 0) {
          mm_new <- model.matrix(~ . - 1, data = df[rows_to_pred, predictors, drop = FALSE])
          final_preds_full[rows_to_pred] <- tryCatch(predict(final_fit, mm_new), error = function(e) rep(NA_real_, length(rows_to_pred)))
        }
      }
    } else if (method == "lgb") {
      mm_train <- model.matrix(~ . - 1, data = full_train_df[, predictors, drop = FALSE])
      dtrain <- lightgbm::lgb.Dataset(data = mm_train, label = full_train_df[[target_var]])
      bp <- best_combo_res$params
      lgb_params <- list(objective = "regression", learning_rate = bp$learning_rate, num_leaves = as.integer(bp$num_leaves))
      final_fit <- tryCatch(lightgbm::lgb.train(params = lgb_params, data = dtrain, nrounds = as.integer(bp$nrounds), verbose = -1), error = function(e) NULL)
      final_preds_full <- rep(NA_real_, nrow(df))
      if (!is.null(final_fit)) {
        rows_to_pred <- which(is.na(df[[target_var]]))
        if (length(rows_to_pred) > 0) {
          mm_new <- model.matrix(~ . - 1, data = df[rows_to_pred, predictors, drop = FALSE])
          final_preds_full[rows_to_pred] <- tryCatch(predict(final_fit, mm_new), error = function(e) rep(NA_real_, length(rows_to_pred)))
        }
      }
    } else if (method == "ranger") {
      bp <- best_combo_res$params
      final_fit <- tryCatch(ranger::ranger(as.formula(paste(target_var, "~", paste(predictors, collapse = " + "))),
                                          data = full_train_df,
                                          num.trees = as.integer(bp$num.trees),
                                          min.node.size = as.integer(bp$min.node.size),
                                          importance = "impurity"), error = function(e) NULL)
      final_preds_full <- rep(NA_real_, nrow(df))
      if (!is.null(final_fit)) {
        rows_to_pred <- which(is.na(df[[target_var]]))
        if (length(rows_to_pred) > 0) {
          final_preds_full[rows_to_pred] <- tryCatch(predict(final_fit, data = df[rows_to_pred, predictors, drop = FALSE])$predictions, error = function(e) rep(NA_real_, length(rows_to_pred)))
        }
      }
    } else {
      final_fit <- NULL
      final_preds_full <- rep(NA_real_, nrow(df))
    }

    all_method_results[[method]] <- list(
      method = method,
      grid = grid,
      combos_res = combos_res,
      best_combo = best_combo_res,
      final_model = final_fit,
      final_imputations = final_preds_full
    )

  } # end methods loop

  # stop parallel plan
  plan(sequential)

  return(list(
    results_by_method = all_method_results,
    params = param_grids,
    meta = list(n_folds = n_folds, n_repeats = n_repeats, mask_frac = mask_frac)
  ))
}

```



```{r}
diagnose_residuals <- function(df_all,
                               bad_prediction,   # tibble with columns index, true, pred, abs_error
                               predictors,
                               p_thresh = 0.1,
                               rf_ntree = 200,
                               rf_min_node_size = 5,
                               seed = 123) {
  library(dplyr); library(broom); library(ggplot2); library(ranger)
  set.seed(seed)

  if (!all(c("index", "true", "pred") %in% names(bad_prediction))) stop("bad_prediction must have index,true,pred")
  bad_idx <- bad_prediction$index
  n_bad <- length(bad_idx)
  n_all <- nrow(df_all)
  other_idx <- setdiff(seq_len(n_all), bad_idx)

  text_lines <- c()
  # Numeric tests
  numeric_vars <- predictors[sapply(df_all[, predictors, drop = FALSE], is.numeric)]
  if (length(numeric_vars) > 0) {
    num_res <- lapply(numeric_vars, function(v) {
      bad_vals <- df_all[[v]][bad_idx]
      good_vals <- df_all[[v]][other_idx]
      # choose t-test if both >= 30 else wilcox
      use_t <- length(na.omit(bad_vals)) >= 30 && length(na.omit(good_vals)) >= 30
      test <- if (use_t) tryCatch(t.test(bad_vals, good_vals), error = function(e) NULL) else tryCatch(wilcox.test(bad_vals, good_vals), error = function(e) NULL)
      tibble::tibble(Variable = v,
                     mean_bad = mean(bad_vals, na.rm = TRUE),
                     mean_rest = mean(good_vals, na.rm = TRUE),
                     diff = mean_bad - mean_rest,
                     p_value = if (!is.null(test)) broom::tidy(test)$p.value else NA_real_)
    })
    num_tbl <- dplyr::bind_rows(num_res)
    top_num <- num_tbl %>% arrange(p_value) %>% head(6)
    for (i in seq_len(nrow(top_num))) {
      text_lines <- c(text_lines, sprintf("- Numeric %s: mean_bad=%.2f, mean_rest=%.2f, Δ=%.2f, p=%.3g",
                                         top_num$Variable[i], top_num$mean_bad[i], top_num$mean_rest[i], top_num$diff[i], top_num$p_value[i]))
    }
  } else text_lines <- c(text_lines, "- No numeric predictors to test.")

  # Factor tests
  factor_vars <- predictors[sapply(df_all[, predictors, drop = FALSE], function(x) is.factor(x) || is.character(x))]
  if (length(factor_vars) > 0) {
    for (v in factor_vars) {
      full_tab <- table(df_all[[v]])
      bad_tab <- table(df_all[[v]][bad_idx])
      all_levels <- union(names(full_tab), names(bad_tab))
      full_tab <- full_tab[all_levels]; bad_tab <- bad_tab[all_levels]
      full_tab[is.na(full_tab)] <- 0; bad_tab[is.na(bad_tab)] <- 0
      p_vals <- sapply(all_levels, function(lvl) {
        bad_yes <- bad_tab[[lvl]]; bad_no <- sum(bad_tab) - bad_yes
        good_yes <- full_tab[[lvl]] - bad_yes; good_no <- sum(full_tab) - full_tab[[lvl]] - bad_no
        tbl <- matrix(c(bad_yes, bad_no, good_yes, good_no), nrow = 2, byrow = TRUE)
        if (any(tbl < 5)) {
          test <- tryCatch(fisher.test(tbl)$p.value, error = function(e) NA_real_)
        } else {
          test <- tryCatch(chisq.test(tbl, correct = FALSE)$p.value, error = function(e) NA_real_)
        }
        test
      })
      over_rep <- (bad_tab / sum(bad_tab)) - (full_tab / sum(full_tab))
      lvl_df <- tibble::tibble(level = all_levels, prop_bad = as.numeric(bad_tab / sum(bad_tab)),
                               prop_all = as.numeric(full_tab / sum(full_tab)), over = as.numeric(over_rep), p = p_vals)
      lvl_df <- lvl_df %>% arrange(p)
      sig <- lvl_df %>% filter(!is.na(p) & p < p_thresh)
      if (nrow(sig) > 0) {
        top <- sig %>% arrange(desc(over)) %>% slice_head(n = 1)
        text_lines <- c(text_lines, sprintf("- Factor %s: level '%s' over-represented (%.2f%% vs %.2f%%; p=%.3g)",
                                           v, top$level, 100*top$prop_bad, 100*top$prop_all, top$p))
      }
    }
  } else text_lines <- c(text_lines, "- No factor predictors to test.")

  # Train RF on residuals (pred - true) restricted to bad indices (if we have residuals)
  resid <- bad_prediction$pred - bad_prediction$true
  rf_plot <- ggplot() + ggtitle("Not enough rows to compute importance")
  if (length(resid) >= 10) {
    rf_df <- df_all[bad_idx, , drop = FALSE]
    rf_df$.resid <- resid
    rf_mod <- ranger::ranger(.resid ~ ., data = rf_df[, c(predictors, ".resid"), drop = FALSE],
                             num.trees = rf_ntree, min.node.size = rf_min_node_size, importance = "impurity")
    imp <- ranger::importance(rf_mod)
    imp_df <- tibble::tibble(variable = names(imp), importance = as.numeric(imp)) %>% arrange(desc(importance))
    rf_plot <- ggplot(imp_df[1:min(30, nrow(imp_df)), , drop = FALSE], aes(x = reorder(variable, importance), y = importance)) +
      geom_col() + coord_flip() + labs(title = "Feature importance for residuals (bad subset)", x = "variable", y = "importance") + theme_minimal()
  }

  text_summary <- if (length(text_lines) > 0) paste(text_lines, collapse = "\n") else "No significant structural differences detected."

  list(text_summary = text_summary, numeric_table = if (exists("num_tbl")) num_tbl else NULL,
       factor_checks = NULL, importance_plot = rf_plot)
}


```


```{r}
# small example param grids (override defaults for speed)
param_grids_custom <- list(
  lm = data.frame(dummy = NA),
  xgb = expand.grid(eta = c(0.05,0.1), max_depth = c(3L), nrounds = c(50L)),
  lgb = expand.grid(learning_rate = c(0.05), num_leaves = c(15L), nrounds = c(50L)),
  ranger = expand.grid(num.trees = c(100L), min.node.size = c(5L))
)

res_cv <- run_imputer_cv(
  df = train_csv,
  target_var = "LotFrontage",
  predictors = setdiff(names(train_csv), "LotFrontage"),
  methods = c("lm","xgb","lgb","ranger"),
  param_grids = param_grids_custom,
  n_folds = 5,
  n_repeats = 3,
  mask_frac = 0.20,
  min_rows_train = 30,
  seed = 123,
  parallel = TRUE   # uses available cores - 1
)

# Inspect results
names(res_cv$results_by_method)
res_cv$results_by_method$xgb$best_combo$rmse   # best combo summary RMSE
res_cv$results_by_method$xgb$final_imputations # final imputations (length == nrow(df)); NA where not predicted


```

```{r}
fe_res <- fe_engineering(
  df = train_csv,
  collapse_factor_min = 10,
  numeric_log = c("LotArea"),
  numeric_sqrt = c("GrLivArea"),
  poly_vars = list(LotArea = 2),
  interactions = list(c("LotArea", "OverallQual")),
  binning = list(LotArea = c(5000, 10000)),
  winsorize = list(LotArea = c(0.01, 0.99)),
  one_hot = FALSE,
  remove_original = FALSE
)

df_fe <- fe_res$data
fe_meta <- fe_res$meta
# Inspect
str(df_fe); print(fe_meta)

```
```{r}
diag_res <- diagnose_residuals(df_all = train_csv,
                               bad_prediction = bad_prediction,
                               predictors = setdiff(names(train_csv), "LotFrontage"),
                               p_thresh = 0.1,
                               rf_ntree = 200)

cat(diag_res$text_summary)
print(diag_res$importance_plot)

```

```{r}
na_count_test
```
Missing data is present in the MSZoning, LotFrontage, Alley, Utilities, Exterior1st, Exterior2nd, MasVnrType, MasVnrArea, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinSF1, BsmtFinType2, BsmtFinSF2, BsmtUnfSF, TotalBsmtSF, BsmtFullBath, BsmtHalfBath, KitchenQual, Functional, FireplaceQu, GarageType, GarageYrBlt, GarageFinish, GarageCars, GarageArea, GarageQual, GarageCond, PoolQC, Fence, MiscFeature, SaleType
For the sake of consistency, we will use broadly the same methods as used in the training dataset to impute the missing values in the test dataset.